---
  title: "hw#4.data624"
  author: "Sang Yoon (Andy) Hwang"
  date: "September 22, 2019"
  output:
  html_document: default
  pdf_document: default
  word_document: default
---
  
# 7.8 Exercises
  
# 7.8.1. Consider the `pigs` series - the number of pigs slaughtered in Victoria each month.
```{r, include=FALSE}
library(fpp2)
library(forecast)
data(pigs)
```

# a. Use the ses() function in R to find the optimal values of alpha and l0 and generate forecasts for the next four months.

From `summary(fc)`, we know that alpha = 0.2971 and l0 = 77260.06. For n the forcasts for the next 4 months, we see that it has a "flat" forecast of 98816.41.
```{r}
fc <- ses(pigs, h = 4)
fc
#fitted(fc)
summary(fc)
```

# b. Compute a 95% prediction interval for the first forecast using  y_^±1.96s where s is the standard deviation of the residuals. Compare your interval with the interval produced by R.

From manual calculation, we know Lo 95 = 78679.97 and Hi 95 = 118952.8 for Sep 1995.
From ses output, we know Lo 95 = 78611.59 and Hi 95 = 119021.2 for Sep 1995.
```{r}

lo95 = fc$mean[1] - (sd(fc$residuals) * 1.96)
hi95 = fc$mean[1] + (sd(fc$residuals) * 1.96)

lo95
hi95
fc

# plot the data with fitted values to forcast
autoplot(fc) + autolayer(fc$fitted)
```

# 7.8.3 Modify your function from the previous exercise to return the sum of squared errors rather than the forecast of the next observation. Then use the `optim()` function to find the optimal values of alpha and l0. Do you get the same values as the `ses()` function?

The results are almost the same but not exactly the same. I think this is due to the scale difference in alpha and l0 when it comes to calculation.
```{r}
# create ses function manually to test forecasted value
ses_manual <- function(y, alpha, l0){
  y_hat <- l0
  for(index in 1:length(y)){
   y_hat <- alpha*y[index] + (1 - alpha)*y_hat 
  }
  cat("Forecast result by ses_optim function: ",
      as.character(y_hat),
      sep = "\n")
}

# Let's compare the forecasting results from built-in SES function in R to manual function using optim
a <- fc$model$par[1]
l0 <- fc$model$par[2]

# They are the same!
ses_manual(pigs, alpha = a, l0 = l0)
cat("Forecast result by built-in ses function: ", fc$mean[1], sep = "\n")

# now, let's change the above function to get alpha and l0
ses_param <- function(pars = c(alpha, l0), y){
  error <- 0
  SSE <- 0
  alpha <- pars[1]
  l0 <- pars[2]
  y_hat <- l0
  
  for(index in 1:length(y)){
    error <- y[index] - y_hat
    SSE <- SSE + error^2
    
    y_hat <- alpha*y[index] + (1 - alpha)*y_hat 
  }
  
  return(SSE)
}

# Let's compare the results for alphan and l0 from built-in SES function in R to manual function using optim
optim_param <- optim(par = c(0.5, pigs[1]), y = pigs, fn = ses_param)

# Almost same
cat("Parameters estimate by manual ses_param function: ", optim_param$par[1], optim_param$par[2], sep = "\n")
cat("Parameters estimate by built-in ses function: ", fc$model$par[1], fc$model$par[2])
```
