---
title: 'DATA 624: Group 2 Homework'
author: 'Juliann McEachern'
output: 
  html_document:
    theme: readable
    highlight: pygments
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

# Assignment 4 

Assignment 4 includes problem 7.1 and 7.3 from the HA text. The following R packages have been used for completion of all homework assignments to date:

```{r dependencies, echo=T, warning=F, message=F, comment=F}
#Textbook Packages
library(fpp2)
library(AppliedPredictiveModeling)
library(mlbench)

#Processing
library(tidyverse)

#Graphing
library(ggplot2)
library(gridExtra)

#Math
library(caret)
library(randomForest)
library(seasonal)
library(psych)
library(corrplot)

#Formatting
library(knitr)
library(kableExtra)
```






## 7.1:  Consider the `pigs` series -- the number of pigs slaughtered in Victoria each month. 
### (a). Use the `ses()` function in R to find the optimal values of $\alpha$ and $\ell_0$, and generate forecasts for the next four months. {.tabset}

#### Forecast Summary

The forecast summary below shows us the optimal values of $\alpha$ and $\ell_0$, as estimated by the `ses` function in R. 

```{r ha-7.1a1, tidy=F}
pigsdata <- window(pigs, start=1980)
fc <- ses(pigsdata, h=4)
summary(fc)
```

#### Visualization

```{r ha-7.1a2, fig.width=10}
autoplot(fc) +
  autolayer(fitted(fc), series="Fitted") +
  ylab("Pigs") + xlab("Year")
```

### (b). Compute a 95% prediction interval for the first forecast using $\hat{y}\pm1.96s$ where $s$ is the standard deviation of the residuals. Compare your interval with the interval produced by R. 

Formula calculations of the confidence interval were obtained using $\bar{x} \pm 1.96s$ and compared to the calculations from the `ses` function in R. 

```{r ha-7.1b1}
# Formula Calculatioin
lower <- fc$mean[1]- 1.96*sd(fc$residuals)
upper <- fc$mean[1]+ 1.96*sd(fc$residuals)

# R Calculation
lowerR <- as.numeric(fc$lower[1, "95%"])
upperR <- as.numeric(fc$upper[1, "95%"])
``` 

```{r ha-7.1b2, echo=F}
lower <- cbind("95%"="Lower","Computed" = lower, "R" = lowerR)
upper <- cbind("95%"="Upper","Computed" = upper, "R" = upperR)
as.data.frame(rbind(lower, upper)) %>% kable(caption="Comparison of 95% Confidence Interval Calculations") %>% kable_styling(c("striped", "hover"))
```

## 7.3: Modify your function from the previous exercise to return the sum of squared errors rather than the forecast of the next observation. Then use the `optim()` function to find the optimal values of $\alpha$ and  $\ell_0$. Do you get the same values as the `ses()` function? {.tabset}

### Function 

The prior question asks us to write a function to implement simple exponential smoothing. I experienced difficulty in doing so, thus turning to [stackexchange](https://stats.stackexchange.com/questions/379727/how-to-use-optim-function-in-r-on-my-custom-residual-sum-of-squares-function) for reference. The functions below are modeled from this source  

```{r ha-7.3a}
# Custom SES Function
custom_ses = function(ts, alpha, l0){
  N = length(ts)
  y_hat = c(l0, 0 * (2:N))
  for (j in 1:(N-1))
    {
    new = alpha * ts[j] + (1-alpha) * y_hat[j]
    y_hat[j+1] = new
  }
  return(y_hat)
}

# SSE function
custom_sse = function(fn, ts){
  alpha <- fn[1]
  l0 <- fn[2]
  N = length(ts)
  fit = custom_ses(ts, alpha, l0)
  res = ts - fit
  out = sum(res^2)/(N-1)
  return(out)
}
```

### Optimization

The `optim` function optimizes the parameters $\alpha$ and $\ell_0$ by using the initial values in the `par` argument to maximize output from the vector defined in our custom functions. 

```{r ha-7.3b1}
# `optim` Function 
optimize <- optim(par = c(0.1, 10000), fn = custom_sse, ts = pigsdata)

# Optimized Parameters
alpha = optimize$par[1]
l0 = optimize$par[2]

# SES Model Parameters 
ses_alpha <-fc$model$fit$par[1]
ses_l0 <-fc$model$fit$par[2]
```

The calculated differences between the R estimated paremeters and optimized paremeters from the custom function were minimal, and can be viewed below: 

```{r ha-7.3b2, echo=F}
lower <- cbind("SES Function"="Custom Optimized","alpha" = alpha, "l0" = l0)
upper <- cbind("SES"="R Estimated","alpha" = ses_alpha, "l0" = ses_l0)
as.data.frame(rbind(lower, upper)) %>% kable(caption="Comparison of Parameter Values") %>% kable_styling(c("striped", "hover"))
```
### Visualization

A timeseries plot of our fitted (estimated) and custom (optimized) values show vast similiarities. There difference between the two is hard to distinguish on the plot. 

```{r, fig.width=10}
fit = custom_ses(pigsdata, alpha = alpha, l0 = l0)
ts_fit <- ts(fit, start=1980, frequency = 12)

autoplot(fc) +
  autolayer(fitted(fc), series="Fitted", size=1.5) +
  autolayer(ts_fit, series="Optimized", linetype = "dashed", size=1.5) +
  ylab("Pigs") + xlab("Year")
```



