# Part A: Forecasting ATM Withdrawals {-#part-a}

```{r library-A, echo=F, message=F, warning=F, error=F, comment=F}
# processing
library(readxl)

# graphs
library(ggplot2)

# formatting
library(default)
library(knitr)
library(kableExtra)
```

```{r settings-A, echo=F, message=F, warning=F, error=F, comment=F}
# Set default augments for code chunks
knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, comment=F, fig.width=10, fig.height = 3)

# Set default augments for `kable_styling()` 
default(kable) <- list("latex")
default(kable_styling)  <- list(latex_options = c("hold_position", "striped"))

# Set default for ggplot theme
default(theme) <- list(axis.text.x = element_text(angle = 90, hjust = 1),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank(),
                       plot.title = element_text(color="#B85231", size=10, face="bold"),
                       plot.subtitle = (element_text(size=8, color="#000000")),
                       legend.title = (element_text(size=10, color="#000000", face="bold")),
                       strip.background = element_rect(color="#000000", 
                                                       fill="#F5E8E4", size=1, linetype="solid"),
                       strip.text.x = element_text(size = 8, color = "#000000", face="bold"))

# GGplot Palette
default(scale_color_brewer) <- list(palette = "OrRd")
```

>  **Instructions:** In part A, I want you to forecast how much cash is taken out of 4 different ATM machines for May 2010.  The data is given in a single file.  The variable `Cash` is provided in hundreds of dollars, other than that it is straight forward.  I am being somewhat ambiguous on purpose.  I am giving you data, please provide your written report on your findings, visuals, discussion and your R code all within a Word readable document, except the forecast which you will put in an Excel readable file.  I must be able to cut and paste your R code and run it in R studio.  Your report must be professional - most of all - readable, EASY to follow.  Let me know what you are thinking, assumptions you are making!  Your forecast is a simple CSV or Excel file that MATCHES the format of the data I provide.

## Exploration {-#a-exploration}

```{r}
atm_data <- read_excel("data/ATM624Data.xlsx") 
```

Data exploration.

## Forecast {-#a-forecast}

Data forecast. 

## Discussion {-#a-discussion}

Discuss findings. 