---
title: "Team 2 - Homework Two"
author: "Vinicio Haro"
date: "DATE"
output:
  pdf_document: default
  html_document:
    df_print: paged
subtitle: 'Assignment 3: KJ 8.1-8.3; KJ 8.7'
---

```{r instructions, echo=F, fig.height=5}
# README: GROUP TWO GUIDELINES
# MEAT&POTATOES:
    # Submissions should be completed in a timely manner, within group internal deadlines. 
    # Thoughtful feedback to all homework submissions must be provided in order to compile work. 
    # Responses to all questions should be answered thoroughly with explanations. 
    # Responses should be proofed and spell checked (F7 shortcut in R) upon completion. 
    # Insert all R libraries used in the library code chunk.
    # Only call plotting and formatting libraries as needed in the RMD to compile assignment 
# FORMATTING
    # UPDATE HOMEWORK YAML WITH NAME AND DATE COMPLETED ONLY 
    # UNIVERSAL LATEX FORMATTING WILL BE APPLIED TO THE FINAL SUBMISSION TO ENSURE EVERYONE                               CAN COMPILE DOCUMENT ON THEIR MACHINE
    # EACH DOCUMENT SHOULD BE KNITTED TO A PDF FOR EACH GROUP MEMBER TO REVIEW.
    # EVERYONE IS INDIVIDUALLY RESPONSIBLE FOR ENSURING THE FILE KNITS PROPERLY. 
    # DEFAULT FORMATTING HAS BEEN SET WITHIN EACH TEMPLATE.  
    # TABLES: 
        # All table outputs should be wrapped using the default knitr and kable_styling settings:                             `%>% kable() %>% kable_styling() %>% row_spec()`
        # Add captions to table where appropriate: `kable(caption="CAPTION")`
    # PLOTS:
        # `fig.height` in code chunk options (see above) should be adjusted to larger size when needed (default=3)
        #  All plots should be done using ggplots 
            # Lables should be used to appropriately when not included default graph:                                             `+labs(title="", subtitle="", x="", y="")`
            # All plots should call `+theme_bw()+theme()` to apply default settings
```

## Dependencies 

```{r, echo = F, message=F, warning=F, error=F, comment=NA, self.contained = F}
# SOURCE DEFAULT SETTINGS
#source('~/GitHub/CUNY_DATA_624/Homework-Two/defaults.R')
source('C:/Users/traveler/Documents/defaults.R')   
options(tinytex.verbose = TRUE)
```

```{r libraries, echo=T}
# Forecast libraries
libraries('mlbench', 'AppliedPredictiveModeling', 'party','gbm', 'Cubist')
# Regression libraries
libraries('randomForest', 'caret', 'rpart','impute','partykit')
# Formatting Libraries
libraries('default', 'knitr', 'kableExtra', 'dplyr', 'sqldf')
# Plotting Libraries
libraries('ggplot2', 'grid', 'ggfortify')
```


## (1) Kuhn & Johnson 8.1

> Recreate the simulated data from Exercise 7.2:

```{r kj-8.1, echo=T}
set.seed(200)
simulated <- mlbench.friedman1(200, sd = 1) 
simulated <- cbind(simulated$x, simulated$y)
simulated <- as.data.frame(simulated) 
colnames(simulated)[ncol(simulated)] <- "y"
```

>> (a). Fit a random forest model to all of the predictors, then estimate the variable importance scores. Did the random forest model significantly use the uninformative predictors (V6-V10)?

The code to the RF model has been provided to us through the literature.

What is the code actually doing? We should dive into the theory. The improtance is calculated with the following formula:

$$
ni_j=W_{left(j)}C_{left(j)}-W_{right(j)}C_{right(j)}
$$

Lets deconstruct the theory. $ni_j$ stands for node importance. $w_j$ is the weighted number of samples reaching node j. $c_j$ is the impurity value of node j. Left(j) is the child node from left split on node j and right(j) is the child node from right split on node j. Once $ni_j$ is calculated,the importance for each variable on a decision tree is calculated with formula I. Formula I is then normalized as shown as formula II.The final feature importance is the average over all trees. We simply find the sum of the features importance value and then divide by all trees T, shown in formula III.

$$
I) \  fi_j \ =\frac{ (\sum_{j \ node \ split \ on \ i}) ni_j }{ ( \sum_{all \ nodes}) ni_k}\\
$$

$$
II) \  \parallel  fi_j  \parallel = \frac{(fi_j)}{\sum_{all \ features}fi_j} \\
$$
$$
III) \ RFfi_j= \frac{ ( \sum_{all \ trees})\parallel  fi_j \parallel}{T}
$$



```{r kj-8.1a}
rf_mod <- randomForest(y ~ ., data = simulated, 
                       importance = TRUE, 
                       ntree = 1000)
rf_imp <- varImp(rf_mod, scale = FALSE)
```

```{r kj-8.1a2}
#dt <- as.matrix(rf_imp)
rf_imp%>% kable(caption="Random Forest Variable Importance") %>% kable_styling() %>% row_spec()
```

Did the random forest model significantly use the uninformative predictors (V6 – V10)?
The most important feature was V1 with an importance of over 8 followed by V2. Variables V6 through V10 are not as significant as v1-V5. 

>> (b). Now add an additional predictor that is highly correlated with one of the informative predictors. Fit another random forest model to these data. Did the importance score for V1 change? What happens when you add another predictor that is also highly correlated with V1? For example:

```{r kj-8.1b-ex, eval=F}
simulated$duplicate1 <- simulated$V1 + rnorm(200) * .1
cor(simulated$duplicate1, simulated$V1) %>% kable(caption="Correlation of New Predictor") %>% kable_styling() %>% row_spec()
```

```{r kj-8.1b}
#code
rf_mod2 <- randomForest(y ~ ., data = simulated, 
  importance = TRUE,
 ntree = 1000)
rf_imp2 <- varImp(rf_mod2, scale = FALSE)

#dt <- as.matrix(rf_imp2)

#dt <- as.matrix(rf_imp2)

#kable(dt)

rf_imp2%>% kable(caption="Random Forest Variable Importance With New Variable") %>% kable_styling() %>% row_spec()
```

By adding a highly correlated variable, we see that V1 importance drops by roughly 2 measures. 

>> (c). Use the `cforest` function in the party package to fit a random forest model using conditional inference trees. The party package function `varimp` can calculate predictor importance. The `conditional` argument of that function toggles between the traditional importance measure and the modified version described in Strobl et al. (2007). Do these importances show the same pattern as the traditional random forest model?

```{r kj-8.1c}
#code

rf_mod3 <- cforest(y ~ ., data=simulated)

#rfdt3 <-as.data.frame(as.matrix(varimp(rf_mod3)))
rfdt3 <-varimp(rf_mod3)

rfdt3 %>% kable(caption="Unconditional CForest Model: Variable Importance") %>% kable_styling() %>% row_spec()
```

```{r kj-8.1c2}
rfdt3b <-varimp(rf_mod3, conditional=T)

rfdt3b  %>% kable(caption="Conditional CForest Model: Variable Importance") %>% kable_styling() %>% row_spec()
```

We see that adding a conditional or non conditional clause for cForest does not have an effect on the importance of variables V1-V6. We some changes in variable importance closer to the upper bound. What is going on in the background?
When we set conditional to true,we compute feature importance by creating permutations built on covariates associated with features of interest. The clause has a backend threshold where it only grabs features with a compliment of the p value greater than the threshold (determined on the backend). This generated score is analagous to beta coefficients for regression models. 

>> (d). Repeat this process with different tree models, such as boosted trees and Cubist. Does the same pattern occur?

Let's try a gradient boosting tree. Why does this tree matter? It is computationaly fast. On a high level, the GB tree starts by taking the average across the target label. We then compute residuals where the residual is simply the difference between the actual value and the predicted value. A decision tree is built on these residuals where each leaf node will contain some prediction along with the residual value. There are some cases where there are more residuals than leaves but on the backend, the GB tree function will compute the average residuals in these cases. We then of course predict new target values and compute new residuals. This process can be repeated until convergece or a tree limit is defined. 

```{r kj-8.1d}
#code
gbm_mod <- gbm(y ~ ., data=simulated, distribution='gaussian')

summary(gbm_mod) %>% kable(caption="Gradient Boosting Tree") %>% kable_styling() %>% row_spec()
```

The summary returns the variable name along with a measure of influence on the target variable. From our GBM tree, we can see v4 is the most influential. V1 and correlated duplicate1 are also much more influential. V6-V10 does not break the top half of our list of influential variables.

The next model we want to try is the cubist model. The cubist model is a rather unique variation on trees. Each leaf in the tree contains a linear regression model. Every layer in the tree alters the predicitons used within each leaf contained model. In other words, the selection of predictors in leaf n is based on the previous splis. It should be noted that each intermediate step between leafs also contain a linear model. Predictions are made via linear models on the on the terminal node.

```{r kj-8.1d2}
#code
cubist_mod <- cubist(x=simulated[,-(ncol(simulated)-1)], y=simulated$y, committees=100)

#dt3<-as.data.frame(as.matrix(varImp(cubist_mod)))
dt3<-varImp(cubist_mod)

dt3 %>% kable(caption="Cubist Model Regression Variable Importance") %>% kable_styling() %>% row_spec()

```

V2 is the most important variable closely followed by V1. We still do not see variables V6-V10 increasing in importance.

## (2) Kuhn & Johnson 8.2 
> Use a simulation to show tree bias with different granularities.
We need to understand the purpose of this problem. The literature covers some key elements when it comes to tree bias based off certain conditions. We know that if we have a variable with many distinct values, then it is said that this variable has a low variance. These types of variables are magnetic to the response variable vs other predictors that have a higher variance. This greatly affects the splits in any tree system and in some cases, splits are done on noise. Typically we try to avoid cases where noise takes the role of the root node.

Let's create a dataset where our predictors will vary in terms of unique values.We then run rpart.
```{r kj-8.2}
#code
set.seed(100)
X1 <- rep(1:2,each=100)
X2 <- rnorm(200,mean=0,sd=2)
#X3 <- sample(0:100 / 100, 200, replace = T)
Y <- X1 + rnorm(200,mean=0,sd=4)
sim <- data.frame(Y=Y, X1=X1, X2=X2)

rpart_mod <- rpart(Y ~ ., data = sim)

#dt4<-as.data.frame(as.matrix(varImp(rpart_mod)))
dt4<-varImp(rpart_mod)

dt4 %>% kable(caption="Tree Bias") %>% kable_styling() %>% row_spec()
```

## (3) Kuhn & Johnson 8.3
> In stochastic gradient boosting the bagging fraction and learning rate will govern the construction of the trees as they are guided by the gradient. Although the optimal values of these parameters should be obtained through the tuning process, it is helpful to understand how the magnitudes of these parameters affect magnitudes of variable importance. Figure 8.24 provides the variable importance plots for boosting using two extreme values for the bagging fraction (0.1 and 0.9) and the learning rate (0.1 and 0.9) for the solubility data. The left-hand plot has both parameters set to 0.1, and the right-hand plot has both set to 0.9:
 
>> (a). Why does the model on the right focus its importance on just the first few of predictors, whereas the model on the left spreads importance across more predictors? 
There are some mechanics that occur on the right hand model that we should point out. First, we see that there is a relationship between the learning rate and how "greedy" the model gets. Recall that a greedy model tries to find the optimal point at each iteration. The relationship states that as the learning rate approaches one, the overall "greed" of the model increases. As a consequence of increased "greed" , the model tends to no longer identify as many variables related to the label variable. We should also address the bagging fraction.  When the bagging fraction increases, we tend to see more usage of data in the model. An analogy is if you consider data to be the fuel of the model (engine),then more of it is consumed as the bagging fraction increases.This also reduces the number of predictors considered to be important. 

>> (b). Which model do you think would be more predictive of other samples? 
We need to consider how altering the parameter values will affect our models. If we were to increase the values in the parameters, we should expect to see model performance decrease, therefore it is safe to say that the left hand model will be more predictive of our other samples. 

>> (c). How would increasing interaction depth affect the slope of predictor importance for either model in Fig.8.24?
The literature states that there is a relationship between interaction depth and feature importance. As we see tree depth increase, we expect a more uniform spread of variable importance across all variables. Visually, this makes the lines bigger in the importance figures shown in 8.2.4.

## (4) Kuhn & Johnson 8.7

>Refer to Exercises 6.3 and 7.5 which describe a chemical manufacturing process. Use the same data imputation, data splitting, and pre-processing steps as before and train several tree-based models:
>> (a). Which tree-based regression model gives the optimal resampling and test set performance? 

```{r kj-8.7a}
#code
data(ChemicalManufacturingProcess)
cd<-ChemicalManufacturingProcess



predictors <- subset(cd,select= -Yield)
yield <- subset(cd,select="Yield")

set.seed(517)
trainingRows <- createDataPartition(yield$Yield, p = 0.8 , list = FALSE)

x_train <- predictors[trainingRows,]
y_train <- yield[trainingRows,]

x_test <- predictors[-trainingRows,]
y_test <- yield[-trainingRows,]

#Pre-process trainPredictors and apply to trainPredictors and testPredictors
pp <- preProcess(x_train,method=c("BoxCox","center","scale","knnImpute"))
transform_x_train <- predict(pp,x_train)
transform_x_test <- predict(pp,x_test)

#Identify and remove NZV
nz_predictors <- nearZeroVar(transform_x_train)
transform_x_train <- transform_x_train[-nz_predictors]
transform_x_test <- transform_x_test[-nz_predictors]

#Identify and remove highly correlated predictors
pred_corr = cor(transform_x_train)
high_corr <- findCorrelation(pred_corr)
transform_x_train <-transform_x_train[, -high_corr]
transform_x_test <- transform_x_test[, -high_corr]

#Set-up trainControl boot method (CV takes way too long)
set.seed(517)
tune_param <- trainControl(method = "boot", number = 25)
```

Build Models 
We will build a standard model for CART, Random Forest, GBM, and Cubist. We already looked at the background of some of the models featured here.

```{r kj-8.7a2}
#CART
set.seed(614)
rpart_param <- expand.grid(maxdepth= seq(1,10,by=1))
rpart_tune <- train(x = transform_x_train, y = y_train ,method= "rpart2",metric = "Rsquared",tuneGrid = rpart_param,trControl = tune_param)

#RF
set.seed(614)
rf_param <- expand.grid(mtry=seq(2,38,by=3))
rf_tune <- train(x = transform_x_train, y =y_train,method = "rf",tuneGrid = rf_param,metric = "Rsquared",
importance = TRUE, trControl = tune_param)

#GBM
set.seed(614)
gbm_param <- expand.grid(interaction.depth=seq(1,6,by=1),n.trees=c(25,50,100,200),shrinkage=c(0.01,0.05,0.1,0.2),n.minobsinnode=c(5, 10, 15))
gbm_tune <- train(x = transform_x_train, y = y_train, method = "gbm",metric = "Rsquared",verbose = FALSE,
tuneGrid = gbm_param,trControl = tune_param)

#cubist
set.seed(614)
cubist_param <- expand.grid(committees = c(1, 5, 10, 20, 50, 100), neighbors = c(0, 1, 3, 5, 7))
cubist_tune <- train(x = transform_x_train, y = y_train,method = "cubist", verbose = FALSE,metric = "Rsquared",
tuneGrid = cubist_param,trControl = tune_param)
```


>> (b). Which predictors are most important in the optimal tree-based regression model? Do either the biological or process variables dominate the list? How do the top 10 important predictors compare to the top 10 predictors from the optimal linear and nonlinear models?

```{r kj-8.7b}

#code
rpart_metric<-as.data.frame(round(rpart_tune$results$Rsquared[best(rpart_tune$results, "Rsquared", maximize = TRUE)],2))
colnames(rpart_metric) <- "r_squared"
row.names(rpart_metric) <- "CART"

rf_metric<-as.data.frame(round(rf_tune$results$Rsquared[best(rf_tune$results, "Rsquared", maximize = TRUE)],2))
colnames(rf_metric) <- "r_squared"
row.names(rf_metric) <- "RF"

gbm_metric<-as.data.frame(round(gbm_tune$results$Rsquared[best(gbm_tune$results, "Rsquared", maximize = TRUE)],2))
colnames(gbm_metric) <- "r_squared"
row.names(gbm_metric) <- "GBM"

cubist_metric<-as.data.frame(round(cubist_tune$results$Rsquared[best(cubist_tune$results, "Rsquared", maximize = TRUE)],2))
colnames(cubist_metric) <- "r_squared"
row.names(cubist_metric) <- "Cubist"

library(sqldf)
metrics<-sqldf("
  select r_squared, 'CART' as model from rpart_metric
  union all 
  select r_squared, 'RF' as model from rf_metric
  union all 
  select r_squared, 'GBM' as model from gbm_metric
  union all 
  select r_squared, 'Cubist' as model from cubist_metric
               ")

metrics %>% kable(caption="R Squared Comparisons") %>% kable_styling() %>% row_spec()
```

Our initial performance inspection should start by checking the variability captured in the data. We generate the r squared value for each model and can observe that the cubist model is the highest performing with 70 percent of the data variability captured in the data. CART captures the least amount of data variability out of our four models. 

```{r kj-8.7b1}
#code
rpart_pred <- predict(rpart_tune, newdata=transform_x_test)
rpart_res <- data.frame()
rpart_res <- rbind(rpart_res, t(postResample(pred=rpart_pred, obs=y_test)))

rf_pred <- predict(rf_tune, newdata=transform_x_test)
rf_res <- data.frame()
rf_res <- rbind(rf_res, t(postResample(pred=rf_pred, obs=y_test)))

gbm_pred <- predict(gbm_tune, newdata=transform_x_test)
gbm_res <- data.frame()
gbm_res <- rbind(gbm_res, t(postResample(pred=gbm_pred, obs=y_test)))

cubist_pred <- predict(cubist_tune, newdata=transform_x_test)
cubist_res <- data.frame()
cubist_res <- rbind(cubist_res, t(postResample(pred=cubist_pred, obs=y_test)))

test_metric<-sqldf(
  "
  select 'CART' as model, RMSE, Rsquared, MAE from rpart_res
  union all
  select 'RF' as model, RMSE, Rsquared, MAE from rf_res
  union all 
  select 'GBM' as model, RMSE, Rsquared, MAE from gbm_res
  union all 
  select 'Cubist' as model, RMSE, Rsquared, MAE from cubist_res
  ")

test_metric %>% kable(caption="R Squared Comparisons (Test)") %>% kable_styling() %>% row_spec()
```

We also examine model performance on the test data. We want to see the predictive performance of each model. Based on performance agaist the test data, Cubist model still has the highest RSquared in addition to the lowest RMSE and MAE. 
I believe ths cubist model outperforms any of the other models from chapter 6 and 7. 

Now we test variable importance for our optimal model. 

```{r kj-8.7b2}
#code
cubist_imp <- varImp(cubist_tune, scale = FALSE)
plot(cubist_imp, top=20, scales = list(y = list(cex = 0.8)))
```


>> (c). Plot the optimal single tree with the distribution of yield in the terminal nodes. Does this view of the data provide additional knowledge about the biological or process predictors and their relationship with yield?

```{r kj-8.7c}
#code
plot(as.party(rpart_tune$finalModel),gp=gpar(fontsize=11))
```

CART is our most optimal single tree. We can use partykit to visualuze our tree. Manufacturing Process 32 becomes our root node. Accordig to our tree, when MP 32 is greater, this results in an increased yield. 


Manufacturing processes 32 and 13 are at the top, with higher values of process 32 being associated with larger yields.  Lower values of process 32 are associated with smaller yields.  However, a lower values of process 32 may be counter-acted with a corresponding lower value of process 13.

## R Code

```{r 03-code, eval=F, echo=T}
# insert code here
# (8.1a)
rf_mod <- randomForest(y ~ ., data = simulated, 
                       importance = TRUE, 
                       ntree = 1000)
rf_imp <- varImp(rf_mod, scale = FALSE)

dt <- as.matrix(rf_imp)

# (8.1b)
#code
rf_mod2 <- randomForest(y ~ ., data = simulated, 
  importance = TRUE,
 ntree = 1000)
rf_imp2 <- varImp(rf_mod2, scale = FALSE)

dt <- as.matrix(rf_imp2)

dt <- as.matrix(rf_imp2)


# (8.1c)
#code
library(party)

rf_mod3 <- cforest(y ~ ., data=simulated)

# (8.1d)
#code
gbm_mod <- gbm(y ~ ., data=simulated, distribution='gaussian')

# (8.2)
#code
set.seed(100)
X1 <- rep(1:2,each=100)
X2 <- rnorm(200,mean=0,sd=2)
#X3 <- sample(0:100 / 100, 200, replace = T)
Y <- X1 + rnorm(200,mean=0,sd=4)
sim <- data.frame(Y=Y, X1=X1, X2=X2)

rpart_mod <- rpart(Y ~ ., data = sim)

dt4<-as.data.frame(as.matrix(varImp(rpart_mod)))

# (8.3a)
# (8.3b)
# (8.3c)

# (8.7a)
#code
data(ChemicalManufacturingProcess)
cd<-ChemicalManufacturingProcess



predictors <- subset(cd,select= -Yield)
yield <- subset(cd,select="Yield")

set.seed(517)
trainingRows <- createDataPartition(yield$Yield, p = 0.8 , list = FALSE)

x_train <- predictors[trainingRows,]
y_train <- yield[trainingRows,]

x_test <- predictors[-trainingRows,]
y_test <- yield[-trainingRows,]

#Pre-process trainPredictors and apply to trainPredictors and testPredictors
pp <- preProcess(x_train,method=c("BoxCox","center","scale","knnImpute"))
transform_x_train <- predict(pp,x_train)
transform_x_test <- predict(pp,x_test)

#Identify and remove NZV
nz_predictors <- nearZeroVar(transform_x_train)
transform_x_train <- transform_x_train[-nz_predictors]
transform_x_test <- transform_x_test[-nz_predictors]

#Identify and remove highly correlated predictors
pred_corr = cor(transform_x_train)
high_corr <- findCorrelation(pred_corr)
transform_x_train <-transform_x_train[, -high_corr]
transform_x_test <- transform_x_test[, -high_corr]

#Set-up trainControl boot method (CV takes way too long)
set.seed(517)
tune_param <- trainControl(method = "boot", number = 25)

#CART
set.seed(614)
rpart_param <- expand.grid(maxdepth= seq(1,10,by=1))
rpart_tune <- train(x = transform_x_train, y = y_train,method = "rpart2",metric = "Rsquared",tuneGrid = rpartGrid,trControl = tune_param)

#RF
set.seed(614)
rf_param <- expand.grid(mtry=seq(2,38,by=3))
rf_tune <- train(x = transform_x_train, y =y_train,method = "rf",tuneGrid = rf_param,metric = "Rsquared",
importance = TRUE, trControl = tune_param)

#GBM
set.seed(614)
gbm_param <- expand.grid(interaction.depth=seq(1,6,by=1),n.trees=c(25,50,100,200),shrinkage=c(0.01,0.05,0.1,0.2),n.minobsinnode=c(5, 10, 15))
gbm_tune <- train(x = transform_x_train, y = y_train, method = "gbm",metric = "Rsquared",verbose = FALSE,
tuneGrid = gbm_param,trControl = tune_param)

#cubist
set.seed(614)
cubist_param <- expand.grid(committees = c(1, 5, 10, 20, 50, 100), neighbors = c(0, 1, 3, 5, 7))
cubist_tune <- train(x = transform_x_train, y = y_train,method = "cubist", verbose = FALSE,metric = "Rsquared",
tuneGrid = cubist_param,trControl = tune_param)

# (8.7b)
#code
rpart_metric<-as.data.frame(round(rpart_tune$results$Rsquared[best(rpart_tune$results, "Rsquared", maximize = TRUE)],2))
colnames(rpart_metric) <- "r_squared"
row.names(rpart_metric) <- "CART"

rf_metric<-as.data.frame(round(rf_tune$results$Rsquared[best(rf_tune$results, "Rsquared", maximize = TRUE)],2))
colnames(rf_metric) <- "r_squared"
row.names(rf_metric) <- "RF"

gbm_metric<-as.data.frame(round(gbm_tune$results$Rsquared[best(gbm_tune$results, "Rsquared", maximize = TRUE)],2))
colnames(gbm_metric) <- "r_squared"
row.names(gbm_metric) <- "GBM"

cubist_metric<-as.data.frame(round(cubist_tune$results$Rsquared[best(cubist_tune$results, "Rsquared", maximize = TRUE)],2))
colnames(cubist_metric) <- "r_squared"
row.names(cubist_metric) <- "Cubist"

library(sqldf)
metrics<-sqldf("
  select r_squared, 'CART' as model from rpart_metric
  union all 
  select r_squared, 'RF' as model from rf_metric
  union all 
  select r_squared, 'GBM' as model from gbm_metric
  union all 
  select r_squared, 'Cubist' as model from cubist_metric
               ")

metrics %>% kable(caption="R Squared Comparisons") %>% kable_styling() %>% row_spec()

#code
rpart_pred <- predict(rpart_tune, newdata=transform_x_test)
rpart_res <- data.frame()
rpart_res <- rbind(rpart_res, t(postResample(pred=rpart_pred, obs=y_test)))

rf_pred <- predict(rf_tune, newdata=transform_x_test)
rf_res <- data.frame()
rf_res <- rbind(rf_res, t(postResample(pred=rf_pred, obs=y_test)))

gbm_pred <- predict(gbm_tune, newdata=transform_x_test)
gbm_res <- data.frame()
gbm_res <- rbind(gbm_res, t(postResample(pred=gbm_pred, obs=y_test)))

cubist_pred <- predict(cubist_tune, newdata=transform_x_test)
cubist_res <- data.frame()
cubist_res <- rbind(cubist_res, t(postResample(pred=cubist_pred, obs=y_test)))

test_metric<-sqldf(
  "
  select 'CART' as model, RMSE, Rsquared, MAE from rpart_res
  union all
  select 'RF' as model, RMSE, Rsquared, MAE from rf_res
  union all 
  select 'GBM' as model, RMSE, Rsquared, MAE from gbm_res
  union all 
  select 'Cubist' as model, RMSE, Rsquared, MAE from cubist_res
  ")
#code
cubist_imp <- varImp(cubist_tune, scale = FALSE)
plot(cubist_imp, top=20, scales = list(y = list(cex = 0.8)))

# (8.7c)
library(partykit)
plot(as.party(rpart_tune$finalModel),gp=gpar(fontsize=11))
```
