# Assignment Four 

*  Hyndman 7.1     
*  Hyndman 7.3   


```{r library-04, echo=F, warning=F, message=F, comment=F, error=F}
#Textbook Packages
library(fpp2)
library(AppliedPredictiveModeling)
library(mlbench)

#Processing
library(tidyverse)

#Graphing
library(ggplot2)
library(gridExtra)

#Math
library(caret)
library(randomForest)
library(seasonal)
library(psych)


#Formatting
library(knitr)
require(kableExtra)
require(default)
```


```{r settings-04, echo=F}
# Set default augments for code chunks
knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, comment=F, fig.width=10, fig.height = 3)

# Set default augments for `kable_styling()` 
#default(kable) <- list(format = "latex", label=NULL)
#default(kable_styling)  <- list(full_width = T,
                                #position="center",
                               # font_size = 10, 
                                #latex_options = c("HOLD_position", "striped"))

# Set default for ggplot theme
default(theme) <- list(axis.text.x = element_text(angle = 90, hjust = 1),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank(),
                       plot.title = element_text(color="#3d248a", size=10, face="bold"),
                       plot.subtitle = (element_text(size=8, color="#000000")),
                       legend.title = (element_text(size=10, color="#000000", face="bold")),
                       strip.background = element_rect(color="#000000", 
                                                       fill="#D3D3D3", size=1, linetype="solid"),
                       strip.text.x = element_text(size = 8, color = "#000000", face="bold"))

# wrap text
if(!knitr:::is_latex_output())
{
  options("width"=56)
  knitr::opts_chunk$set(tidy.opts=list(width.cutoff=56, indent = 2), tidy = TRUE)
  knitr::opts_chunk$set(fig.pos = 'H')
}
```



## Hyndman 7.1   

> Consider the `pigs` series -- the number of pigs slaughtered in Victoria each month. 

### a. Use the `ses()` function in R to find the optimal values of $\alpha$ and $\ell_0$, and generate forecasts for the next four months.

#### Forecast Summary

The forecast summary below shows us the optimal values of $\alpha$ and $\ell_0$, as estimated by the `ses` function in R. 

```{r ha-7.1a1, tidy=F}
pigsdata <- window(pigs, start=1980)
fc <- ses(pigsdata, h=4)
summary(fc) 
```

#### Visualization

```{r ha-7.1a2}
autoplot(fc) +
  autolayer(fitted(fc), series="Fitted") +
  ylab("Pigs") + xlab("Year")+theme_bw()+theme()
```

### b. Compute a 95% prediction interval for the first forecast using $\hat{y}\pm1.96s$ where $s$ is the standard deviation of the residuals. Compare your interval with the interval produced by R.

Formula calculations of the confidence interval were obtained using $\bar{x} \pm 1.96s$ and compared to the calculations from the `ses` function in R. 

```{r ha-7.1b1}
# Formula Calculatioin
lower <- fc$mean[1]- 1.96*sd(fc$residuals)
upper <- fc$mean[1]+ 1.96*sd(fc$residuals)

# R Calculation
lowerR <- as.numeric(fc$lower[1, "95%"])
upperR <- as.numeric(fc$upper[1, "95%"])

# Table Output
lower <- cbind("95_perct"="Lower","Computed" = lower, "R" = lowerR)
upper <- cbind("95_perct"="Upper","Computed" = upper, "R" = upperR)
as.data.frame(rbind(lower, upper)) %>% 
  kable("latex") %>% 
  kable_styling(latex_options = "hold_position")
``` 


## Hyndman 7.3   

> Modify your function from the previous exercise to return the sum of squared errors rather than the forecast of the next observation. Then use the `optim()` function to find the optimal values of $\alpha$ and  $\ell_0$. Do you get the same values as the `ses()` function? 

#### Function

The prior question asks us to write a function to implement simple exponential smoothing. As this was a challenge, we made reference to  [stackexchange](https://stats.stackexchange.com/questions/379727/how-to-use-optim-function-in-r-on-my-custom-residual-sum-of-squares-function) while crafting what follows.  

```{r ha-7.3a}
# Custom SES Function
custom_ses = function(ts, alpha, l0){
  N = length(ts)
  y_hat = c(l0, 0 * (2:N))
  for (j in 1:(N-1))
    {
    new = alpha * ts[j] + (1-alpha) * y_hat[j]
    y_hat[j+1] = new
  }
  return(y_hat)
}

# SSE function
custom_sse = function(fn, ts){
  alpha <- fn[1]
  l0 <- fn[2]
  N = length(ts)
  fit = custom_ses(ts, alpha, l0)
  res = ts - fit
  out = sum(res^2)/(N-1)
  return(out)
}
```

#### Optimization

The `optim` function optimizes the parameters $\alpha$ and $\ell_0$ by using the initial values in the `par` argument to maximize output from the vector defined in our custom functions. 

```{r ha-7.3b1}
# `optim` Function 
optimize <- optim(par = c(0.1, 10000), fn = custom_sse, ts = pigsdata)

# Optimized Parameters
alpha = optimize$par[1]
l0 = optimize$par[2]

# SES Model Parameters 
ses_alpha <-fc$model$fit$par[1]
ses_l0 <-fc$model$fit$par[2]
```

The calculated differences between the R estimated paremeters and optimized parameters from the custom function were minimal, as is visible in the comparison below below: 

```{r table2}
# Table Output
lower <- cbind("SES Function"="Custom Optimized","alpha" = alpha, "l0" = l0)
upper <- cbind("SES"="R Estimated","alpha" = ses_alpha, "l0" = ses_l0)
as.data.frame(rbind(lower, upper)) %>% 
  kable("latex") %>% 
  kable_styling(latex_options = c("hold_position", "striped")) 
```

#### Visualization

A timeseries plot of our fitted (estimated) and custom (optimized) values are very similar, such that the difference between the two is difficult to discern in the plot. 

```{r  ha-7.3b3, fig.width=10}
fit = custom_ses(pigsdata, alpha = alpha, l0 = l0)
ts_fit <- ts(fit, start=1980, frequency = 12)

autoplot(fc) +
  autolayer(fitted(fc), series="Fitted", size=1.5) +
  autolayer(ts_fit, series="Optimized", linetype = "dashed", size=1.5) +
  ylab("Pigs") + xlab("Year")+theme_bw()+theme()
```
