# Assignment Five

*  Hyndman 7.5 
*  Hyndman 7.6 
*  Hyndman 7.10

```{r library-05, echo=T, warning=F, message=F, comment=F}
#Textbook Packages
library(fpp2)
library(AppliedPredictiveModeling)
library(mlbench)

#Processing
library(tidyverse)

#Graphing
library(ggplot2)
library(grid)
library(gridExtra)
library(lemon)

#Math
library(caret)
library(forecast)
library(randomForest)
library(seasonal)
library(psych)
library(corrplot)

#Formatting
library(knitr)
library(kableExtra)
library(default)
```

```{r settings-05, echo=F}
# Set default augments for code chunks
knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, comment=F, fig.width=10, fig.height = 3)

# Set default for ggplot theme
default(theme) <- list(axis.text.x = element_text(angle = 90, hjust = 1),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank(),
                       plot.title = element_text(color="#3d248a", size=10, face="bold"),
                       plot.subtitle = (element_text(size=8, color="#000000")),
                       legend.title = (element_text(size=10, color="#000000", face="bold")),
                       strip.background = element_rect(color="#000000", 
                                                       fill="#D3D3D3", size=1, linetype="solid"),
                       strip.text.x = element_text(size = 8, color = "#000000", face="bold"))

# wrap text
if(!knitr:::is_latex_output())
{
  options("width"=45)
  knitr::opts_chunk$set(tidy.opts=list(width.cutoff=40, indent = 2), tidy = TRUE)
  knitr::opts_chunk$set(fig.pos = 'H')
}
```

## Hyndman 7.5 

>  Data set books contains the daily sales of paperback and hardcover books at the same store. The task is to forecast the next four days' sales for paperback and hardcover books. 

### a. Plot the series and discuss the main features of the data. 

#### Timeseries Plot

The `books` data shows an overall positive trend over the 30-day period for paperback and hardcover books. There is heavy fluctuation between days, suggesting seasonality within the time series. While paperbacks and hardcovers both have a similar range of unit sales, peaks and valleys are not well synchronized.

```{r ha-7.5a1}
autoplot(books, facet=TRUE) + 
  labs(title="Daily Sales of Paperback and Hardcover Books", 
       x="Days", 
       y="Sales") + 
  theme_bw() + theme()
``` 

#### Decomposed Plot

We can look at the decomposed timeseries to better understand the trend and seasonality on a weekly basis. Sales tend to peak three times per week and increase throughout the entire month. The paperback books trend increases more drastically between weeks two and four, while hardcover books appear to have smoother incline. Hardcover books also show sharper fluctuations in the seasonal component. 

```{r ha-7.5a2, fig.height= 5}
paperback <- books[,1]
hardcover <- books[,2] 

pb_plot <- ts(paperback, frequency=7) %>% 
  decompose(type="multiplicative") %>% 
  autoplot() +
  labs(title = "Paperback Books",
       x = "Weeks") + theme_bw()+ theme()

hc_plot <- ts(hardcover, frequency = 7) %>% 
  decompose(type="multiplicative") %>% 
  autoplot() +
  labs(title = "Hardcover Books",
       x = "Weeks") + theme_bw()+ theme() 

grid.arrange(pb_plot, 
             hc_plot, 
             ncol = 2, 
             top=textGrob("Decomposition of Multiplicative TS",
                          gp = gpar(fontface = "bold", cex = 1.5))) 
``` 

### b. Use the `ses()` function to forecast each series, and plot the forecasts. {.tabset}

#### Paperback SES

```{r ha-7.5b1}
pb_ses <- ses(paperback, h=4)
summary(pb_ses)
``` 

#### Hardcover SES

```{r ha-7.5b2}
hc_ses <- ses(hardcover, h=4)
summary(hc_ses)
``` 

#### Plot

```{r ha-7.5b3}
pb_ses_plot <- autoplot(pb_ses, size=1, fcol="#0044cc") +
  autolayer(fitted(pb_ses), series="Fitted", alpha=.75, size=1.5, color="#99b3ff") +
  labs(title="Paperback Books", y="Sales", x="Days")+ theme_bw()+ theme() 

hc_ses_plot <- autoplot(hc_ses, size=1, fcol="#0044cc") +
  autolayer(fitted(hc_ses), series="Fitted", alpha=.75, size=1.5, color="#99b3ff") +
  labs(title="Hardcover Books", y="Sales", x="Days")+ theme_bw() + theme() 

grid.arrange(pb_ses_plot, hc_ses_plot, ncol = 2, 
             top=textGrob("4-Day Forecast (SES)",
                          gp = gpar(fontface = "bold", cex = 1.5)))
``` 

As SES may work well when their is no clear trend or seasonal pattern. As it generates the next probable value and uses that for all future forecasting values, however `ses()` may not be the best long-term forecasting model for this particular data.

### c. Compute the RMSE values for the training data in each case.

The root mean squared error (RMSE) is calculated below and tells us the spread between the residual errors from our predicted and observed using the SES method. The RMSE for paperback books is slightly higher than hardcover, suggesting better accuracy in the later's predictions.

```{r ha-7.5c1, eval=F}
#manual calculation
sqrt(mean(pb_ses$residuals^2))
sqrt(mean(hc_ses$residuals^2))

#caret package
RMSE(fitted(pb_ses), paperback)
RMSE(fitted(hc_ses), hardcover)

``` 

```{r ha-7.5c2, echo=F}

pb_rmse_ses <- RMSE(fitted(pb_ses), paperback)
hc_rmse_ses <- RMSE(fitted(hc_ses), hardcover)

# as.data.frame(cbind("paperback" = pb_rmse_ses, "hardcover" = hc_rmse_ses)) %>% 
#   kable(caption="RMSE Values for Book Sales Forecast", "latex") %>% 
#   kable_styling(c("striped", "hold_position"))
```   

**Paperback RMSE:** `r pb_rmse_ses`   
**Hardcover RMSE:** `r hc_rmse_ses`  

## Hyndman 7.6 

### a. Now apply Holt’s linear method to the paperback and hardback series and compute four-day forecasts in each case.

```{r ha7.6a, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE}  
hardcover <- ts(books[,2], start = 1, frequency = 7)
softbound <- ts(books[,1], start = 1, frequency = 7)

hb <- ses(hardcover, h=5)
sb <- ses(softbound, h=5)

acc_soft <- accuracy(sb)
acc_hard <-accuracy(hb)

holt_soft <- holt(softbound, seasonal = 'additive', h=4)
holt_hard <- holt(hardcover, seasonal = 'additive', h=4)

acc_holt_hard <-accuracy(holt_hard)
acc_holt_soft <-accuracy(holt_soft)

pb_holt_plot <- autoplot(holt_soft, size=1, fcol="#7733ff") +
  autolayer(fitted(holt_soft), series="Fitted", alpha=.75, size=1.5, color="#e0b3ff") +
  labs(title="Paperback Books", y="Sales", x="Days") + theme_bw()+ theme() 

hc_holt_plot <- autoplot(holt_hard, size=1, fcol="#7733ff") +
  autolayer(fitted(holt_hard), series="Fitted", alpha=.75, size=1.5, color="#e0b3ff") +
  labs(title="Hardcover Books", y="Sales", x="Days")+ theme_bw()+ theme() 

grid.arrange(pb_holt_plot, hc_holt_plot, ncol = 2, 
             top=textGrob("4-Day Forecast (Holt)",
                          gp = gpar(fontface = "bold", cex = 1.5)))
```

As would be expected, Holt's linear method generates a linear trend forecast.

### b. Compare the RMSE measures of Holt’s method for the two series to those of simple exponential smoothing in the previous question. Discuss the merits of the two forecasting methods for these data sets.

> Remember that Holt’s method is using one more parameter than SES.

**RMSE Exponential Smoothed Hardcover Sales:** `r round(acc_hard[2],3)`  
**RMSE Exponential Smoothed Paperback Sales:** `r round(acc_soft[2],3)`  
**RMSE Holt-Winters Hardcover Sales:**  `r round(acc_holt_hard[2],3)`  
**RMSE Holt-Winters Paperback Sales:**  `r round(acc_holt_soft[2],3)`  

The Holt-Winter method does indeed reduce the root mean squared errors for both the paperback and hardcover book sales. This would suggest that there is in fact some level of trend acting on the book sales series.

### c. Compare the forecasts for the two series using both methods. Which do you think is best?

```{r ha7.6c}  
fc_hh<-forecast(holt_hard)
fc_sh<-forecast(holt_soft)

fc_hes<-forecast(hb, h=4)
fc_ses<-forecast(sb, h=4)

p1<-autoplot(softbound) +
    autolayer(fc_ses, series = 'SES Paperback', PI=FALSE) +
    autolayer(fc_sh, series = 'Holt Paperback', PI=FALSE) +
        labs(title ='Paperback Sales',
         x = "Year",
         y = "Daily Sales of Paperback Books")+theme_bw()+theme(legend.position = "bottom")


p2<-autoplot(hardcover) +
    autolayer(fc_hes, series = 'SES Paperback', PI=FALSE) +
    autolayer(fc_hh, series = 'Holt Paperback', PI=FALSE) +
        labs(title ='Hardcover Sales',
         x = "Year",
         y = "Daily Sales of Hardcover Books") +theme_bw()+ theme(legend.position = "bottom")

grid.arrange(p1,p2, nrow=1, 
             top=textGrob("Methods Comparison",
                          gp = gpar(fontface = "bold", cex = 1.5)))
```


While Holt's method provided better accuracy measures (RMSE), it is a tough choice to pick one model over the other because neither is particularly robust over the long term. the SES is providing a continuous flat prediction which we can see is clearly not how the series has evolved in the past. On the other hand Holt's method does account for the trend but not the seasonality, so at some point it too would fail. Choosing a model from the two should be done based on your end need (ses being more conservative than holt's method) and with the understanding that more than a few observations forward and either of them will likely underperform.

#### d. Calculate a 95% prediction interval for the first forecast for each series, using the RMSE values and assuming normal errors. Compare your intervals with those produced using ses and holt.

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE}  

pi_ses_s<-1.96*acc_soft[2]
pi_ses_h<-1.96*acc_hard[2]

pi_hl_s<-1.96*acc_holt_soft[2]
pi_hl_s<-1.96*acc_holt_hard[2]

l_interval_ses_s <- round(fc_ses$mean[1]-pi_ses_s,3)
l_interval_ses_h <- round(fc_hes$mean[1]-pi_ses_h,3)

l_interval_hl_s <- round(fc_sh$mean[1]-pi_hl_s,3)
l_interval_hl_h <- round(fc_hh$mean[1]-pi_hl_s,3)

u_interval_ses_s <- round(fc_ses$mean[1]+pi_ses_s,3)
u_interval_ses_h <- round(fc_hes$mean[1]+pi_ses_h,3)

u_interval_hl_s <- round(fc_sh$mean[1]+pi_hl_s,3)
u_interval_hl_h <- round(fc_hh$mean[1]+pi_hl_s,3)


```

|                | Model Lower | Model Upper | Calculated Lower | Calculated Upper |
|----------------|-------------|-------------|------------------|------------------|
| SES Softbound  | `r round(fc_ses$lower[1,2],3)`        | `r round(fc_ses$upper[1,2],3)`        | `r l_interval_ses_s`             | `r u_interval_ses_s`             |
| SES Hardbound  | `r round(fc_hes$lower[1,2],3)`        | `r round(fc_hes$upper[1,2],3)`        | `r l_interval_ses_h`             | `r u_interval_ses_h`             |
| Holt Softbound | `r round(fc_sh$lower[1,2],3)`        | `r round(fc_sh$upper[1,2],3)`        | `r l_interval_hl_s`             | `r u_interval_hl_s`             |
| Holt Hardbound | `r round(fc_hh$lower[1,2],3)`        | `r round(fc_hh$upper[1,2],3)`        | `r l_interval_hl_h `             | `r u_interval_hl_h`             |

The manually calculated and library-generated prediction intervals are not exactly the same, but they are close. Manual intervals are a bit less consistent with the library computed interval values using the Holt's linear method than they are with exponentially smoothed (SES) version.  Additionally, as SES entails a flat forecast with higher RMSE, its interval is slightly wider than that generated by Holt's linear method.

## Hyndman 7.10 

> For this exercise use data set ukcars, the quarterly UK passenger vehicle production data from 1977Q1–2005Q1.  

### a. Plot the data and describe the main features of the series.  

```{r ha-7.10a, fig.height=3, fig.width=8}
ukcdata <- window(ukcars, start=1977, end=2005)
autoplot(ukcdata) +
  labs(title = "Quarterly UK Passenger Vehicle Production Data", y="Production", x="Year")
``` 

There appears to be a general upward trend in production from 1982-Q3 to 2000-Q1. The overall time series shows fluxuations between years suggesting an underlying seasonality component. There also is some less structured periodicity and random-like variation. 


### b. Decompose the series using STL and obtain the seasonally adjusted data.

### STL Decomposition

Through decomposition, we can confirm the overall increasing trend described in part (a). There are consistant seasonal fluxations which can we can futher evaluate using a seasonal plot. 

```{r ha7.10b1, fig.height= 5}
ukc_stl <- stl(ukcars, t.window = 13, s.window = 'periodic', robust = TRUE)

ukc_stl %>%
    autoplot()+
    ggtitle("Decomposition of UK Cars usign 'STL' ")+
  theme_bw()+theme()
```  

### Seasonal Plots 

With the seasonal plot, we can see that our overal seasonality component increases between Q1-Q2, decreases Q2-Q3, and increases Q3-Q4. The intensity of the fluxuations vary by year. There are a few exceptions to this observation, for example, 2000 shows a two quarter decrease between Q1-Q3. 

The subseries plot further confirms our seasonal pattern and shows that the smallest pattern changes occur between Q1-Q2. The decreases and increases between Q2-Q4 are much more distingushable.

```{r ha-7.10b2, fig.height=4, fig.width=8}
ukc_s_plot <- ggseasonplot(ukcdata, year.labels=TRUE, year.labels.left=T, col=rainbow(12,s = 0.6, v = 0.75)) +
  labs(title="Seasonal plot", y="Production") +
  theme_bw()+theme()

ukc_ss_plot <- ggsubseriesplot(ukcdata) +
  labs(title="Subseries plot", y="Production") +theme_bw()+theme()

grid.arrange(ukc_s_plot, 
             ukc_ss_plot, 
             ncol = 2,
             top="Quarterly UK Passenger Vehicle Production Data")
                           
```

We can capture the seasonal component observed in the data and control for the changes overtime using a seasonal adjustment. We compared the seasonally adjusted data to the observed data for `ukcars` below: 

```{r ha-7.10b3, fig.height=3, fig.width=8}
ukc_seas <- seasadj(ukc_stl)

autoplot(ukcdata, series="Data") +
  autolayer(ukc_seas, series="Seasonally Adjusted") +
   labs(title = "Quarterly UK Passenger Vehicle Production Data", 
       subtitle = "Seasonally Adjusted Comparison", 
       x = "Year", 
       y = "Production")+
  scale_colour_manual(values=c("#bfbfbf","#99b3ff"),
                      breaks=c("Data","Seasonally Adjusted","Trend")) +
  theme_bw()+
  theme(legend.position="bottom") 
```


### c. Forecast the next two years of the series using an additive damped trend method applied to the seasonally adjusted data. (This can be done in one step using stlf() with arguments etsmodel="AAN", damped=TRUE.)

As the data is seasonally adjusted, we use an AAN model (error = A, trend = A, and seasonal = N). Damped models are often better at forecasting long-run values.

#### Forecast

```{r ha-7.10c1}
ukc_stlf1 <- stlf(ukcdata, s.window = "periodic", etsmodel="AAN", damped=TRUE, h=8)
rmse_holt_d <- accuracy(ukc_stlf1)[2]

kable(forecast(ukc_stlf1), "latex")%>%
  kable_styling(latex_options = c("hold_position", "striped"))
``` 

#### Plot 

```{r ha-7.10c2, fig.width=8, fig.height=3}
autoplot(ukc_stlf1, size=1.25, fcol="#0044cc") +
  autolayer(fitted(ukc_stlf1), series="Fitted", alpha=.5, size=1.5, color="#99b3ff") +
  labs(y="Production", x="Years")+theme_bw()+theme()
```


### d. Forecast the next two years of the series using Holt’s linear method applied to the seasonally adjusted data (as before but with damped=FALSE).  

We again use an AAN model (error = A, trend = A, seasonality = N) - this time without damping, so point forecasts are a little bit higher.

#### Forecast

```{r ha-7.10d1}
ukc_stlf2 <- stlf(ukcdata, s.window = "periodic", etsmodel="AAN", damped=FALSE, h=8)
rmse_holt_u <- accuracy(ukc_stlf2)[2]

kable(forecast(ukc_stlf2), "latex")%>%
  kable_styling(latex_options = c("hold_position", "striped"))
``` 

#### Plot

```{r ha-7.10d2, fig.width=8, fig.height=3}
autoplot(ukc_stlf2, size=1.25, fcol="#0044cc") +
  autolayer(fitted(ukc_stlf2), series="Fitted", alpha=.5, size=1.5, color="#99b3ff") +
  labs(y="Production", x="Years")+theme_bw()+theme()
```

### e. Now use ets() to choose a seasonal model for the data.  

ETS automatically chooses the best model by minimizing AIC, which results in a error = A, trend = N, seasonal = A model.

#### Seasonal Model

```{r ha7.10e, fig.height= 5}
ukc_ets <- ets(ukcdata)
summary(ukc_ets)
ukc_ets_fc <- ukc_ets %>% forecast(h=8)
rmse_ets <- accuracy(ukc_ets)[2]
```  

#### Forecast

```{r}
kable(ukc_ets_fc, "latex")%>%
  kable_styling(latex_options = c("hold_position", "striped"))
```


#### Plot

```{r ha-7.10e2, fig.width=8, fig.height=3}
autoplot(ukc_ets_fc, size=1.25, fcol="#0044cc") +
  autolayer(fitted(ukc_ets), series="Fitted", alpha=.5, size=1.5, color="#99b3ff") +
  labs(y="Production", x="Years")
```



### f.  Compare the RMSE of the ETS model with the RMSE of the models you obtained using STL decompositions. Which gives the better in-sample fits?    

Our STL decomponsition, using additive damped trend method with no seasonal components ($A, A_d, N$), produced the best in-sample fit. This model provided us with the lowest RMSE, suggesting the highest degree of accuracy between our predicted and observed data. We should note, however, there is a very small variation in RMSE between all STL and ETS models. 

*  **STL:** `r rmse_holt_u` 
*  **HOLT:** `r rmse_holt_d` 
*  **ETS:** `r rmse_ets`  

### g.  Compare the forecasts from the three approaches? Which seems most reasonable?  

The forecasts from the ETS ($A, N, A$) model appears most reasonable and fits our data best. This is not surprising as the `ets` function automatically selects the methodology based on the provided data. The forecasts using STL were both linear, whereas the ETS model was able to predict seasonal pattern changes across our 2-year (8-quarter) prediction.

```{r ha-7.10g, fig.width=8, fig.height=4.5}
ukc_series <- cbind("Data" = ukcdata, 
                   "Forecast (STL1)" = ukc_stlf1[["mean"]], 
                   "Forecast (STL2)" = ukc_stlf2[["mean"]],
                   "Forecast (ETS)" = ukc_ets_fc[["mean"]])

autoplot(ukc_series, size=1, alpha=.75) + 
  scale_color_manual(name = "", 
                     values=c("#000000", "#0044cc", "#99b3ff", "#e0b3ff"))+
  labs(title = "Quarterly UK Passenger Vehicle Production Data", 
       subtitle = "2-Year Forecast Comparison using STL and ETS Models", 
       x="Years", 
       y="Production") +
  theme_bw()+
  theme(legend.position="bottom") 
``` 

### h. Check the residuals of your preferred model.  

The plots below contains our residuals data from the preferred ETS model. We observed constant variance and normally distribution of our residuals, suggesting our model has adequately captured the information from the `ukcars` data

```{r ha-7.10h}
cbind('Residuals' = residuals(ukc_ets),      
      'Forecast errors' = residuals(ukc_ets, type='innovation')) %>%  
  autoplot(facet=TRUE) + 
  labs(title="Residuals from ETS (A,N,A) model of UKcars data", x="Year")+theme_bw()+theme()
``` 

