---
title: 'DATA 624: Project 2'
author:
- Vinicio Haro
- Sang Yoon (Andy) Hwang
- Julian McEachern
- Jeremy O'Brien
- Bethany Poulin
date: '10 December 2019'
output: 
  pdf_document:
    includes:
      in_header: preamble.tex
    latex_engine: xelatex
    keep_tex: yes
    toc: true
    number_sections: no
documentclass: report
subparagraph: yes
---

File for final submission of Project 2. 

```{r formatting, echo = F, message=F, warning=F, error=F, comment=NA}
source('~/GitHub/CUNY_DATA_624/Project_Two/defaults.R')
```

```{r load-source, echo = F, message=F, warning=F, error=F, comment=NA, results='hide'}
source('~/GitHub/CUNY_DATA_624/Project_Two/script.R')
```


# Model Performance 

*  Set1 = Caret: bagImputed; no additional pre-processing  
*  Set2 = Caret: bagImputed; PreP `method=c('center', 'scale', 'nzv', 'BoxCox')`

#### Train Performance:

```{r}
tbl.perf.train1 %>% 
  kable(caption="Train1 Performance", booktabs=T, digits=4) %>% 
  kable_styling() 

tbl.perf.train2 %>% 
  kable(caption="Train2 Performance", booktabs=T, digits=4) %>% 
  kable_styling() 
```

#### Test Accuracy:

```{r}
tbl.perf.test1 %>% 
  kable(caption="Test1 Performance", booktabs=T, digits=4) %>% 
  kable_styling() 

tbl.perf.test2 %>% 
  kable(caption="Test2 Performance", booktabs=T, digits=4) %>% 
  kable_styling() 

```

# New Model : RPART

Cool graph if we add this or replace RF with it -->

**Metrics**

Train with transformation:

```{r, echo=T}
library(party)
library(partykit)
library(caret)
library(MLmetrics)
#install.packages("rattle")
library(rattle)
set.seed(58677)

grid_rpart <- expand.grid(maxdepth = 1:20)
fit.rpart <- train(PH~., data=train_trans, metric="RMSE", method = "rpart2", tuneGrid = grid_rpart, tuneLength=tl, trControl=trC)

rp.PERF <- rbind(getTrainPerf(fit.rpart))
rp.MAPE <- cbind(MAPE = MAPE(fit.rpart$pred$pred, fit.rpart$pred$obs))
rp.ACC <- cbind(rp.MAPE, rp.PERF)




fit.rpart2 <- train(PH~., data=train_trans, metric="RMSE", method = "rpart", tuneLength=tl, trControl=trC)

rp.PERF2 <- rbind(getTrainPerf(fit.rpart2))
rp.MAPE2 <- cbind(MAPE = MAPE(fit.rpart2$pred$pred, fit.rpart2$pred$obs))
rp.ACC2 <- cbind(rp.MAPE2, rp.PERF2)


rbind(rp.ACC, rp.ACC2) %>% kable() %>% kable_styling()
```

**Tune Grid**

```{r}
p1<-ggplot(fit.rpart) + theme_bw() + theme() + labs(title = "Cart1 CV TUNE GRID")

p2<-ggplot(fit.rpart2) + theme_bw() + theme() + labs(title = "Cart2 CV TUNE GRID")

gridExtra::grid.arrange(p1, p2, nrow=1)

```

**Cool Plot**

```{r, fig.height=4}
fancyRpartPlot(fit.rpart$finalModel,
               main = NA,
               sub = "Visualization of Cart1 Trees Influence on pH",
               caption = NA,
               palettes = "PuBuGn",
               type = 2) 


fancyRpartPlot(fit.rpart2$finalModel,
               main = NA,
               sub = "Visualization of Cart2 Trees Influence on pH",
               caption = NA,
               palettes = "Purples",
               type = 2) 
```

**ALT PLOT**

Do not recomment. Seems much more difficult to customize.

```{r, fig.height=4}
rpartY <- as.party.rpart(fit.rpart$finalModel)
plot.party(rpartY)

rpartY <- as.party.rpart(fit.rpart2$finalModel)
plot.party(rpartY)
```

