---
title: "Team 2 - Homework Two"
subtitle: "Assignment 2: KJ 7.2; KJ 7.5"
author: "Juliann McEachern"
date: "10/23/19"
output: 
  pdf_document
---

## Dependencies 

```{r, echo = F, message=F, warning=F, error=F, comment=NA, self.contained = F}
# SOURCE DEFAULT SETTINGS
source('~/GitHub/CUNY_DATA_624/Homework-Two/defaults.R')
```

```{r libraries, echo=T}
# predictive modeling
libraries('mlbench', 'caret','mice','AppliedPredictiveModeling', 'recipes', 'tibble', 'tidyverse')

# Formatting Libraries
libraries('default', 'knitr', 'kableExtra')

# Plotting Libraries
libraries('ggplot2', 'grid', 'ggfortify')
```

## (1) Kuhn & Johnson 7.2

>  Friedman (1991) introduced several benchmark data sets create by simulation. One of these simulations used the following nonlinear equation to create data: $y = 10\text{sin}(\pi x_1 x_2)+20(x_3-0.5)^2+10x_4+5x_5+N(0,\sigma^2)$; where the $x$ values are random variables uniformly distributed between $[0, 1]$ (there are also 5 other non-informative variables also created in the simulation). 

**The package `mlbench` contains a function called `mlbench.friedman1` that simulates these data:** 

```{r kj-7.2-ex1, echo=T}
set.seed(200) 
trainingData <- mlbench.friedman1(200, sd = 1)
trainingData$x <- data.frame(trainingData$x) 
featurePlot(trainingData$x, trainingData$y) 
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x) 
```

>> **(a) Tune several models on these data.**

```{r kj-7.2-ex2, echo=F}
set.seed(100)

# Example 
knnModel <- train(x = trainingData$x,
                  y = trainingData$y, 
                  method = "knn",
                  preProc = c("center", "scale"), 
                  tuneLength = 10) 
knnPred <- predict(knnModel, newdata = testData$x) 
knnPerf <- postResample(pred = knnPred, obs = testData$y)


# Model 1
lmModel <- train(x = trainingData$x,
                  y = trainingData$y, 
                  method = 'lm',
                  preProc = 'pca', 
                  trControl = trainControl(method = "repeatedcv", repeats = 5))
lmPred <- predict(lmModel, newdata = testData$x) 
lmPerf <- postResample(pred = lmPred, obs = testData$y)

# Model 2
plsModel <- train(x = trainingData$x,
                  y = trainingData$y, 
                  method = 'pls',
                  preProc = 'pca', 
                  tuneLength = 5,
                  trControl = trainControl(method = "repeatedcv", repeats = 5))
plsPred <- predict(plsModel, newdata = testData$x) 
plsPerf <- postResample(pred = plsPred, obs = testData$y)

# Model 3
marsModel <- train(x = trainingData$x,
                 y = trainingData$y, 
                 method = 'earth',
                 tuneGrid = expand.grid(degree=1, nprune=10),
                 trControl = trainControl(method = "repeatedcv", repeats = 5))
marsPred <- predict(marsModel, newdata = testData$x) 
marsPerf <- postResample(pred = marsPred, obs = testData$y)
```

### For example: 

**Train set model & performance:**

```{r kj-7.2-models}
knnModel 
```

**Test set performance values:**

```{r}
knnPerf
```

### Model 1: 

**Train set model & performance:**

```{r kj-7.2-1a}
lmModel 
```

**Test set performance values:**

```{r kj-7.2-1b}
lmPerf 
```

### Model 2: 

**Train set model & performance:**
```{r kj-7.2-2a}
plsModel 
```

**Test set performance values:**
```{r kj-7.2-2b}
plsPerf 
```

### Model 3: 

**Train set model & performance:**
```{r kj-7.2-3-a}
marsModel 
```

**Test set performance values:**
```{r kj-7.2-3b}
marsPerf 
```

>> **(b) Which models appear to give the best performance? Does MARS select the informative predictors (those named X1-X5)?**

The MARS model has the lowest RMSE accuracy scores for both our training and test sets. This model appeared to give the best performance. 

```{r kj-7.2-4a}
performance_table <- rbind("knnTrain"=c("RMSE"=knnModel$results$RMSE[6],"RSquared"=knnModel$results$RMSE[6],"MAE"=knnModel$results$RMSE[6]),"knnTest"=knnPerf, "lmTrain"=c("RMSE"=lmModel$results$RMSE,"RSquared"=lmModel$results$Rsquared,"MAE"=lmModel$results$MAE), "lmTest"=lmPerf, "plsTrain"=c(plsModel$results$RMSE[2],plsModel$results$Rsquared[2],plsModel$results$MAE[2]),"plsTest"=plsPerf,"marsTrain"=c(marsModel$results$RMSE,marsModel$results$Rsquared,marsModel$results$MAE), "marsTest"=marsPerf) %>% kable(caption="Model Performance", digits=4) %>% kable_styling() %>% row_spec() %>% row_spec(row=7:8, background ="#d9f2e6")

performance_table
```

In addition, the MARS model selected the important indicator variables: X1-X5.

```{r kj-7.2-4b}
marsImp <- varImp(marsModel)

marsImp$importance %>% kable(caption="MARS Model - Variable Importance", digits=2) %>% kable_styling()
```

## (2) Kuhn & Johnson 7.5

>  Exercise 6.3 describes data for a chemical manufacturing process. Use the same data imputation, data splitting, and pre-processing steps as before and train several nonlinear regression models.

```{r kj-7.5, results='hide'}
# FOR FINAL HW SUBMISSION, DO NOT REPEAT STEPS: JUST CALL VARIABLES FROM PRIOR ASSIGNMENT.

data(ChemicalManufacturingProcess)
CMP_Impute <- mice(ChemicalManufacturingProcess, m=5, printFlag=F)
CMP_DF <- mice::complete(CMP_Impute, 2)

# Set random seed 
set.seed(1)

# Create Partition for Train/Test Splits
trainingRows <- createDataPartition(CMP_DF$Yield, p = .80, list= FALSE)

# Split Train/Test Data 
train <- CMP_DF[trainingRows, ] 
test <- CMP_DF[-trainingRows, ] 

# Pre-Process Recipe 
rec <- recipes::recipe(CMP_DF, Yield~.)
rec <- rec %>% step_nzv( all_predictors(), options = list(freq_cut = 95/5, unique_cut = 10 ) )
prep_rec = prep(rec, training = CMP_DF)
CMP_DF_TF = bake( prep_rec, CMP_DF)

# Create Partition for Train/Test Splits
trainingRows <- createDataPartition(CMP_DF_TF$Yield, p = .80, list= FALSE)

# Split Train/Test Data 
train <- CMP_DF_TF[trainingRows, ] 
test <- CMP_DF_TF[-trainingRows, ] 

# Non-Linear Model 1: Support Vector Machines with Radial Basis Function Kernel
svmModel <- train(Yield~., data=train, method = "svmRadial", preProcess="pca", trControl = trainControl(method = "cv", number = 5), tuneLength=5)
svmPred <- predict(svmModel, newdata = test) 
svmPerf <- postResample(pred = svmPred, obs = test$Yield)

# Non-Linear Model 2: Bayesian Ridge Regression (Model Averaged) 
bma_rrModel <- train(Yield~., data=train, method = 'blassoAveraged', trControl = trainControl(method = "cv", number = 5), tuneLength=5)
bma_rrPred <- predict(bma_rrModel, newdata = test) 
bma_rrPerf <- postResample(pred = bma_rrPred, obs = test$Yield)

# Non-Linear Model 3: k-Nearest Neighbors (KNN)
knnModel <- train(Yield~., data=train, method = 'knn',   trControl = trainControl(method = "cv", number = 5), tuneLength=5)
knnPred <- predict(knnModel, newdata = test) 
knnPerf <- postResample(pred = knnPred, obs = test$Yield)

# Non-Linear Model 4: Multivariate Adaptive Regression Spline (MARS)
marsModel <- train(Yield~., data=train, method = 'earth', tuneGrid = expand.grid(degree=1, nprune=10), trControl = trainControl(method = "repeatedcv", repeats = 5))
marsPred <- predict(marsModel, newdata = test) 
marsPerf <- postResample(pred = marsPred, obs = test$Yield)
```

>> **(a) Which nonlinear regression model gives the optimal resampling and test set performance? **

We trained four models on the chemical manufacturing process data: Support Vector Machines with Radial Basis Function Kernel (SVM), Bayesian Ridge Regression (Model Averaged), k-Nearest Neighbors (KNN), and Multivariate Adaptive Regression Spline (MARS). 

We found that the SVM and Bayesian Ridge approach produced the lowest train accuracy score. However, he train accuracy for the Baysian Ridge was lower than the test accuracy, indicating that model may have been overfitted to the training data. MARS also produced similiarly low accuracy scores. MARS test RMSE was slightly lower than SVM's test accuracy measures, however SVM outperformed MARS with the training accuracy. There was a smaller difference between the train and test accuracy with the SVM method, thus we choose this as our optimal model for resampling and test set performance.

```{r kj-7.5a}
performance_table2 <- rbind("svmTrain"=c(svmModel$results$RMSE[5],svmModel$results$Rsquared[5],svmModel$results$MAE[5]),"svmTest"=svmPerf,"knnTrain"=c("RMSE"=knnModel$results$RMSE[2],"RSquared"=knnModel$results$RMSE[2],"MAE"=knnModel$results$RMSE[2]),"knnTest"=knnPerf, "bma_rrTrain" = c("RMSE"=bma_rrModel$results$RMSE, "RSquared" = bma_rrModel$results$Rsquared, "MAE"=bma_rrModel$results$MAE), "bma_rrTest"=bma_rrPerf, "marsTrain" = c(marsModel$results$RMSE, marsModel$results$Rsquared, marsModel$results$MAE), "marsTest"=marsPerf) %>% kable(caption="Model Performance", digits = 4) %>% kable_styling() %>% row_spec() %>% row_spec(row=1:2, background ="#d9f2e6")

performance_table2
```

>> **(b) Which predictors are most important in the optimal nonlinear regression model? Do either the biological or process variables dominate the list? How do the top ten important predictors compare to the top ten predictors from the optimal linear model? **

The `caret` package does not allow the `varImp` function to work on SVM models, thus we used MARS to evaluate predictor importance of our optimal nonlinear regression models as these MARS ans SVM had similiar accuracy performance. 

In our homework from Chapter 6, our PLS linear model identified mostly biological process variables as the important predictors. This differs from the non-linear important predictors identified in our MARS model. The MARS method calculated variable importance for 7 indicators, of which 6 were manufacturing variables.

```{r kj-7.5b, fig.height=4}
marsImp <- varImp(marsModel)


mars_most_important <- as.data.frame(marsImp$importance) %>% rownames_to_column("Variable") %>% filter(Overall>0)  

plot2 <- ggplot(mars_most_important, aes(x=reorder(Variable, Overall), y=Overall)) + geom_point() + geom_segment(aes(x=Variable,xend=Variable,y=0,yend=Overall)) + labs(title="Variable Importance", subtitle="SVM Model for CMP Data", x="", y="Importance")+ coord_flip()+theme_bw()+theme()

plot2
```


>> **(c) Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model. Do these plots reveal intuition about the biological or process predictors and their relationship with yield?**

The following shows correlation between our top predictor and response variables using our MARS model.

```{r kj-7.5c-1}
top<-mars_most_important$Variable
cor<-cor(train[, top], train$Yield, method="pearson")
cor %>% as.data.frame()%>%rownames_to_column("Variable")%>%rename("Correlation"=V1)%>%arrange(desc(Correlation))%>%kable(caption="Correlation") %>% kable_styling() 
```

We can use a scatterplot to further look at their relationship. The `manufacturingprocess09` indicator has a more linear relationship with yield, whereas the variables exibit a non-linear pattern when examined against yield.

```{r kj-7.5c-2, fig.height=5}
CMP_DF[, top] %>% cbind("Yield"=CMP_DF$Yield) %>% gather(key="Variable", value="Value", -Yield) %>%
  ggplot(aes(Value, Yield, color=Variable)) +
  geom_point(color="grey") +
  geom_smooth(stat="smooth", color="black", method = "loess") +
  facet_wrap(~Variable, scales='free_x', nrow=2) +
  labs(title="Variable Importance Scatterplot",y="Yield Outcome", x="Raw Material Measurements")+
  theme_bw()+
  theme(legend.position = 'none', axis.text.x = element_text(angle = 45, hjust = 1)) 

```

