# Part B: Forecasting Power {-#part-b}

```{r library-B, echo=F, message=F, warning=F, error=F, comment=F}
# processing
library(readxl)

# graphs
library(ggplot2)

# formatting
library(default)
library(knitr)
library(kableExtra)
```

```{r settings-B, echo=F, message=F, warning=F, error=F, comment=F}
# Set default augments for code chunks
knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, comment=F, fig.width=10, fig.height = 3)

# Set default augments for `kable_styling()` 
default(kable) <- list("latex")
default(kable_styling)  <- list(latex_options = c("hold_position", "striped"))

# Set default for ggplot theme
default(theme) <- list(axis.text.x = element_text(angle = 90, hjust = 1),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank(),
                       plot.title = element_text(color="#B85231", size=10, face="bold"),
                       plot.subtitle = (element_text(size=8, color="#000000")),
                       legend.title = (element_text(size=10, color="#000000", face="bold")),
                       strip.background = element_rect(color="#000000", 
                                                       fill="#F5E8E4", size=1, linetype="solid"),
                       strip.text.x = element_text(size = 8, color = "#000000", face="bold"))

# GGplot Palette
default(scale_color_brewer) <- list(palette = "OrRd")
```

> **Instructions:** Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.  Your assignment is to model these data and a monthly forecast for 2014.  The data is given in a single file.  The variable 'KWH' is power consumption in Kilowatt hours, the rest is straight forward.    Add these to your existing files above - clearly labeled.  

## Data Exploration {-#b-exploration}

Explore data. 

```{r}
power_data <- read_excel("data/ResidentialCustomerForecastLoad-624.xlsx") 
```

## Data Model {-#b-model}

Model data. 

## Forecast {-#b-forecast}

Data forecast. 

## Discussion {-#b-discussion}

Discuss findings. 
