---
title: 'DATA 624: Project 1'
author:
- Vinicio Haro
- Sang Yoon (Andy) Hwang
- Julian McEachern
- Jeremy O'Brien
- Bethany Poulin
date: 'October 22, 2019'
documentclass: book
subparagraph: yes
classoption: openany
output: 
  pdf_document:
    highlight: tango
    includes:
      in_header: preamble.tex
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: yes
    number_sections: true
    toc: yes
    toc_depth: 2
---

# Overview {-#overview}

We split the work into three sections for Project 1. Individual team members each took lead on individual problem. Jermey and Julian focused on Part A, Sang Yoon (Andy) and Vinicio worked on Part B, and Bethany took lead on Part C.  Juliann created an overall format for the assignment to be used and all team members collectively worked together on reviewing and merging our finished product. 

## Dependencies {-#dependencies}

The following R libraries were used to complete this assignment:

```{r getting-started, echo=T, eval=T, message=F, warning=F, error=F, comment=F}
library(easypackages)

libraries('knitr', 'kableExtra', 'default')

# Processing
libraries('readxl', 'tidyverse', 'janitor', 'imputeTS', 'tsoutliers')

# Timeseries 
libraries('urca', 'forecast', 'timetk')

# Graphing
libraries('ggplot2', 'grid', 'gridExtra', 'ggfortify','ggpubr', 'scales')
```

## Data {-#data}

Data was stored within our group repository and imported below using the `readxl` package. Each individual question was solved within an R script and the data was sourced into our main report. For replication purposes, we also made our R scripts available within our appendix. All forecasts were exported and saved a `.csv` file in our [github repository]((https://github.com/JeremyOBrien16/CUNY_DATA_624/tree/master/Project%20One/) folder named forecasts.

```{r, eval=F}
# Data Aquisition
atm_data <- read_excel("data/ATM624Data.xlsx") 
power_data <- read_excel("data/ResidentialCustomerForecastLoad-624.xlsx") 
pipe1_data <- read_excel("data/Waterflow_Pipe1.xlsx")
pipe2_data <- read_excel("data/Waterflow_Pipe2.xlsx")

# Source Code
source('~/GitHub/CUNY_DATA_624/Project One/scripts/Part-A.R')
source('~/GitHub/CUNY_DATA_624/Project One/scripts/Part-B.R')
source('~/GitHub/CUNY_DATA_624/Project One/scripts/Part-C.R')
```

```{r settings-A-JM, echo=F, message=F, warning=F, error=F, comment=F}
### UNIVERSAL DATA SOURCING & DEFAULT SETTINGS FOR PROJECT

library(knitr)
library(kableExtra)
library(default)

# Load All Sourced Code Here >>> 
suppressWarnings(source("scripts/Part-A.R"))
suppressWarnings(source("scripts/Part-B.R"))

# Set default augments for code chunks
knitr::opts_chunk$set(echo = F, message=F, warning=F, error=F, comment=F, fig.width=10, fig.height = 3)

# Set default augments for `kable_styling()` 
default(kable) <- list(format="latex")
default(kable_styling)  <- list(latex_options = c("HOLD_position", "striped"))
default(row_spec) <- list(row=0, bold=T)

# Set default for ggplot theme
default(theme) <- list(axis.text.x = element_text(angle = 0, hjust = NULL),
                       plot.title = element_text(color="#4c4c4c", size=12, face="bold"),
                       plot.subtitle = (element_text(size=8, color="#000000")),
                       legend.title = (element_text(size=10, color="#000000", face="bold")),
                       strip.background = element_rect(color="#000000", 
                                                       fill="#cccdd0", size=.75,linetype="solid"),
                       strip.text.x = element_text(size = 8, color = "#000000", face="bold"))

# GGplot Palette
default(scale_color_brewer) <- list(palette = 'RdPu', direction=1)
```


# Part A: ATMs

>  **Instructions:** In part A, I want you to forecast how much cash is taken out of 4 different ATM machines for May 2010.  The data is given in a single file.  The variable `Cash` is provided in hundreds of dollars, other than that it is straight forward.  I am being somewhat ambiguous on purpose.  I am giving you data, please provide your written report on your findings, visuals, discussion and your R code all within a Word readable document, except the forecast which you will put in an Excel readable file.  I must be able to cut and paste your R code and run it in R studio.  Your report must be professional - most of all - readable, EASY to follow.  Let me know what you are thinking, assumptions you are making!  Your forecast is a simple CSV or Excel file that MATCHES the format of the data I provide.

## Exploration

The data covers a period of Friday May 1, 2010 through Saturday April 30, 2010. While reviewing the data, we identified that the original data file contained `NA` values in our `ATM` and `Cash` columns for 14 observations between May 1 and 14, 2010. As these contain no information, we removed these missing values and transformed the dataset into a wide format. 

```{r, include=F}
#SILENCED CODE CHUNK FOR DISCUSSION
# Box plot hard to view and interpret with large outlier. Suggest either removing outlier or
# deleting plot.

# Examine distribution of values to identify outliers
atm_data %>% 
  group_by(ATM) %>% 
  ggplot(aes(x = ATM, y = Cash,fill=ATM)) +
  geom_boxplot() +
  labs(title="ATM Boxplot",x="", y="Cash (in hundreds)") +
  theme_bw() +
  theme(legend.position = 'none', 
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1)) +
  scale_fill_brewer(palette = 'RdPu')
```

Our initial review also revealed that ATM2 contained one missing value on 2009-10-25 and that ATM4 contained a potential outlier of \$1,123 on 2010-02-09. We replaced both values with the corresponding mean value of each machine. 

We examined summary statistics for each ATM time series:  

+  ATM1 and ATM2 have pretty normal distributions; ATM1's daily mean cash dispensed is \$84, and ATM2's is \$62. 
+  ATM3 only dispensed cash on the last three days of the time series - as this provides few data points on which to forecast, we'll need to treat it specially. 
+  ATM4 has a similar mean to ATM1, but skew and kurtosis suggest the impact of an outlier Wednesday, February 10, 2010.  If this ATM is located in the Northeastern United States, this may have a relationship to a blizzard which struck on that day.  

```{r}
# plot atms as scatterplot
atm %>% 
  # re-gather observations for facet plot
  gather(key=ATM, value=Cash, ATM1,ATM2, ATM3,ATM4) %>% 
  # remove NA value from ATM2
  filter(complete.cases(.)) %>% 
  # plot 
  ggplot(aes(DATE, Cash, color=ATM)) +
  geom_point() +
  geom_smooth(method="loess") +
  facet_wrap(~ATM, scales='free_x', nrow=1) +
  labs(title="ATM Scatterplot",x="", y="Cash (in hundreds)")+
  theme_bw()+
  theme(legend.position = 'none', axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer()
```

Last, we used a scatterplot to examine the correlation between cash withdrawals and dates for each machine.  We identified similiar patterns between ATM1 and ATM4, which show non-linear fluctuations that suggest a potential trend component in these timeseries. ATM2 follows a relatively linear path and decreases overtime. This changes in the last few observations, where withdrawals begin to increase. As mentioned, there are only 3 observed transactions for ATM3 that appear at the end of the captured time period. 

## Timeseries Plots 

Our cleaned dataframe was then converted into a timeseries format. The time series plots show high weekly variance, for ATM1, ATM2, and ATM4 - consistent with our takeaway from the scatterplots. 

These plots also remind us that ATM3 only dispensed cash on 3 days at the end of the timespan, with a daily range between \$82 and \$96.  Given the paucity of observations in the training data, the simplest possible approach to forecasting ATM3, averaging, is likely best.  Given that ATM3 distributed no cash until April 28, 2010, we'll assume that it was not operating until then and only include the three day window of non-zero observations in the forecast.

```{r, fig.height=5}
autoplot(atm_ts, facets = T)+
  labs(title = "Daily ATM Transactions", 
       subtitle = "Series: ATM1", y="Cash (in hundreds)", x="Weeks")+
  theme_bw()+ theme()
```

## Evaluation 

We constructed our initial timeseries for ATM1, ATM2, and ATM4 using a weekly frequency. Our ACF plots for each ATM showcases large, decreasing lags starting at 7. This pattern continues in a multiple of seven, which confirms our assumption about seasonality within the observed data. These lags are indicative of a weekly pattern. 
\newline

```{r fig.height=4}
p1<-ggAcf(ATM1_ts)+ labs(title="ACF: ATM1")+ theme_bw()+theme()
p2<-ggPacf(ATM1_ts)+ labs(title="PACF: ATM1")+ theme_bw()+ theme()
p3<-ggAcf(ATM2_ts)+ labs(title="ACF: ATM2")+ theme_bw()+theme()
p4<-ggPacf(ATM2_ts)+ labs(title="PACF: ATM2")+ theme_bw()+ theme()
p5<-ggAcf(ATM4_ts)+ labs(title="ACF: ATM4")+ theme_bw()+theme()
p6<-ggPacf(ATM4_ts)+ labs(title="PACF: ATM4")+ theme_bw()+ theme()

grid.arrange(grob=p1, p3, p5, p2, p4, p6, ncol=3)
```

Our plots further suggest that the ATM data is non-stationary. We performed a unit root test using the `ur.kpss()` function to confirm this observation. The test results below show that differencing is required on all ATM2 and ATM4 series. ATM1 falls just below the cut-off critical value, but could still benefit from differencing due to the observed seasonal pattern.   

```{r}
urATM1<-cbind("ATM"="ATM1", "No-Diff"=round(ATM1_ur@teststat,4),"Diff-1" =round(ATM1d_ur@teststat,4))

urATM2<-cbind("ATM"="ATM2", "No-Diff"=round(ATM2_ur@teststat,4),"Diff-1" =round(ATM2d_ur@teststat,4))

urATM4<-cbind("ATM"="ATM4", "No-Diff"=round(ATM4_ur@teststat,4),"Diff-1" =round(ATM4d_ur@teststat,4))

rbind(urATM1, urATM2,urATM4) %>% kable(caption="KPSS unit root test") %>% kable_styling() %>% row_spec()
```

### Modeling 

We used `auto.arima()` and set `D=1` to account for seasonal differencing of our data to select the best ARIMA models for ATM1, ATM2, and ATM4. The full models and accuracy statistics for each series can be viewed in the appendix.

*  **ATM1**: ARIMA$(0,0,2)(0,1,1)_7$ 
*  **ATM2**: ARIMA$(2,0,2)(0,1,1)_7$
*  **ATM3**: MEAN
*  **ATM4**: ARIMA$(0,0,2)(0,1,1)_7$ 

The residual ACF plots contain no pattern and the lags fall within the critical value, which suggest they are white noise and not autocorrelated. Further, the residual histograms follow a relatively normal distribution, which confirms that the models adequately fits the observed data. 

```{r}
p1<-ggAcf(ATM1_AA$residuals, lag=21)+ labs(title="ATM1", x="Lag", y="") +theme_bw()+theme()
p2<- ggpubr::gghistogram(ATM1_AA$residuals, fill="peachpuff1")+
  labs(title="ATM1", subtitle="ARIMA(0,0,2)(0,1,1)[7]",x="")+
  theme_bw()+theme()
p3<-ggAcf(ATM2_AA$residuals, lag=21)+ labs(title="ATM2",x="Lag", y="") +theme_bw()+theme()
p4<-ggpubr::gghistogram(ATM2_AA$residuals,  fill="lightpink")+
  labs(title="ATM2", subtitle="ARIMA(2,0,2)(0,1,1)[7]",x="")+
  theme_bw()+theme()
p5<-ggAcf(ATM4_AA$residuals, lag=21)+ labs(title="ATM4",x="Lag", y="") +theme_bw()+theme()
p6<-ggpubr::gghistogram(ATM4_AA$residuals,fill="deeppink4")+
  labs(title="ATM4", subtitle="ARIMA(0,0,2)(0,1,1)[7]", x="")+
  theme_bw()+theme()

grid.arrange(grob=p1, p3, p5, ncol=3, top=textGrob(label="ACF Plots of Residuals"))
             
grid.arrange(p2, p4, p6, ncol=3, top=textGrob(label="Histograms of Residuals"))
```

## Forecast

A forecast for the month of May will be 31 days in length. We applied a forecast to each series for 31 days, which span across 5 weeks, in May 2010. The numeric forecasts can be viewed in a table output in the appendix section and are also located within our data output folder.   

```{r, fig.height=10}

p1<-autoplot(ATM1_AA$fitted)+autolayer(ATM1_fc, color="peachpuff1")+
  coord_cartesian(xlim = c(40, 57.5))+
  labs(title = "ATM1 Series", x="Weeks", y="Cash")+theme_bw()+theme()+
  scale_x_continuous(breaks=seq(1,60,by=3))

p2<-autoplot(ATM2_AA$fitted)+autolayer(ATM1_fc,  color="lightpink")+
  coord_cartesian(xlim = c(40, 57.5))+
  labs(title = "ATM2 Series", x="Weeks", y="Cash")+theme_bw()+theme()+
  scale_x_continuous(breaks=seq(1,60,by=3))

p3<-ATM3_plotdata %>% 
  ggplot()+
  geom_line(aes(x = Date/7, y = Cash))+
  geom_line(aes(x = Date/7, y=`Point Forecast`))+
  geom_ribbon(aes(x = Date/7, ymin = `Lo 95`, ymax = `Hi 95`), linetype = 'blank', fill = 'deeppink4', alpha = .4)+
  geom_ribbon(aes(x = Date/7, ymin = `Lo 80`, ymax = `Hi 80`), linetype = 'blank', fill = 'deeppink4', alpha = .2)+
  coord_cartesian(xlim = c(1, 57.5))+
  labs(title = "ATM3 Series", x="Weeks", y="Cash")+theme_bw()+theme()+
  scale_x_continuous(breaks=seq(1,60,by=3))

p4<-autoplot(ATM4_AA$fitted)+autolayer(ATM1_fc, color="darkorchid4")+
  coord_cartesian(xlim = c(1, 57.5))+
  labs(title = "ATM4 Series", x="Weeks", y="Cash")+theme_bw()+theme()+
  scale_x_continuous(breaks=seq(1,60,by=3))

grid.arrange(p1, p2, p3, p4, ncol=1, top=textGrob(label="ATM Forecasts"))
```


# Part B: Forecasting Power

> **Instructions:** Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.  Your assignment is to model these data and a monthly forecast for 2014.  The data is given in a single file.  The variable 'KWH' is power consumption in Kilowatt hours, the rest is straight forward.    Add these to your existing files above - clearly labeled.  

## Exploration

We observed there was a missing value in September 2008. We used imputation method called na.interpolation which performs a technique in numerical analysis which estimates a value from known data points. For our case, linear method using first order Taylor polynomial is used.

## Time Series Plot 

```{r}
# time series plot
autoplot(ts_data) +
labs(title = "Monthly Residential Power Usage", subtitle = "Time Series: 01/98 - 12/13", y = "KWH (in Millions)")+
theme_bw()+theme()+scale_y_continuous(label = unit_format(unit = "m", scale = 1e-6))
```

Our initial time series plot reveal annual seasonality within this time series. The box plot/seasonality plot actually reveals where power consumption fluctuations occur within each of the cycle positions. We can speculate that this could be due to there being no major Holidays that require power draining decor plus we assume minimal AC usage during the cold months. 

## Evaluation 

We see power consumption increase between the months of June and August. This must be tied to AC usage during the warmer months of a year and finally power usage dips from September to Novemeber with a small spike in December. We speculate that thisis due to transitioning out of summer. The spike in December could be connected to the usage or Holiday lights being kept on. 

Within the overall TS plot, we see a dip in July 2010. This could be due to a power outtage during a hot summer month. This can certainly be considered to be an outlier within this TS. Using TSOutliers, we can actually identify the index where our outliers may be. TSoutliers also replaces the outlier using Box-Cox. If set lambda=auto, then TSoutliers will automatically perform Box-Cox transformation.

The ACF plot shows that autocorrelations are well outside the significant space indicating the series is not white noise, non-stationary. 

```{r, fig.height=6}
# season plot
p1 <- ggseasonplot(ts_data)+
  labs(title="Seasonal Plot")+theme_bw()+theme(legend.position = 'none')+
  scale_y_continuous(label = unit_format(unit = "m", scale = 1e-6))

# sub-season plot
p2 <- ggsubseriesplot(ts_data)+labs(title="Subseries Plot", y="")+theme_bw()+theme()+
  scale_y_continuous(label = unit_format(unit = "m", scale = 1e-6))

# ggAcf
p3 <- ggAcf(ts_data)+labs(title="ACF Plot", y="")+theme_bw()+theme()

# STL decomposition
p4 <- autoplot(stl1)+theme_bw()+theme()+scale_y_continuous(label = unit_format(unit = "m", scale = 1e-6))

grid.arrange(grob=p1, p2, p3, p4, ncol=2,   
             layout_matrix = rbind(c(1, 4),
                                   c(2, 4),
                                   c(3, 4)),
             top=textGrob(label="Diagnostic Plots"))
```

## Data Model 

Out of the models we built, we can make some preliminary observations. The residuals for each of our models does not have a major deviance from normality, however residuals of Model #1: ARIMA do not have an extended number of bins distorting the normality proximity but we can say it is still fairly normally distributed. 
 
The residual ACF plots show residual autocorrelations for each of our models. Model #1: ARIMA has less autocorrelation than the other three models. Model 1 is well within the 95% limits indicated by the dotted blue lines. 
 
If we examine the Ljung-Box test results for our models, the only model with a p-value > 0.05 is Model #1: ARIMA. This implies that the residuals from other models are not independent, hence not white noise. The full model summary can be viewed in the appendix.

### Model #1: ARIMA

```{r warning=FALSE, message=FALSE}
checkresiduals(arima_fc)
```

### Model #2: STL (no-demped) - MNN

```{r}
checkresiduals(stl_ndemp)
```

### Model #2-2: STL (demped) - MAdN

```{r}
checkresiduals(stl_demp)
```

### Model #3: ets - MNM
```{r}
checkresiduals(ets_model)
```

## Forecast

The `auto.arima()` function performs cross validation on hyperparameter tuning to find the best model with parameters  of `order` and `seasonal` that minimize `AIC`. This gave us **arima_model**: ARIMA$(3,0,2)(2,1,0)12$ with drift resulting `AIC` = 5332.24. 

Since ARIMA is the only reliable model, as other models failed Ljung test, we will plot forecasts of ARIMA only. The forecasted values can be viewed in the appendix. 

### Model #1: ARIMA

```{r}
autoplot(arima_fc)+
  scale_y_continuous(label = unit_format(unit = "m", scale = 1e-6))+ 
  labs(title="Forecast: Residential Power - Model 1", subtitle="ARIMA(3,0,2)(2,1,0)[12] with drift", y = "KWH (in Millions)")+theme_bw()+theme()

## unable to get autoplot autolayer working :(
## Error: Invalid input: date_trans works with objects of class Date only

#forecast::autoplot(arima_fc)+forecast::autolayer(arima_auto$fitted, series="Fitted")
#  labs(y = "KWH (in Millions)")+
#  theme_classic()+theme(legend.position = 'bottom')+
#  scale_y_continuous(label = unit_format(unit = "m", scale = 1e-6))
```

## Discussion 

We implemented a cross validation method of testing for `h=12`. The process randomly chooses 12 points to measure and take the average of RMSEs. By definition, a lower RMSE on test set is attributed with a better forecast on unseen data. 

Using Time series cross-validation, we compute RMSE on testset (`h=12`). We would have to pick the model with the lowest RMSE on test set as our final model if we had more than 1 model to compare. In our case, since we only have 1 model left after Ljung test, we  have no choice but to pick seasonal ARIMA model as our final choice. Cross-validation test shows that RMSE on test is around 720k when RMSE on training is around 589k. We can conclude the model is not necessarily overfitted. Given that MAPE on training is less than 7, it is not a suprising result.

```{r}
paste("RMSE - train:",rmse_train_arima) 
paste("RMSE - test:",rmse_test_arima)
```



# Part C: Waterflow

>  **Instructions:** Part C consists of two data sets.  These are simple 2 columns sets, however they have different time stamps.  Your optional assignment is to time-base sequence the data and aggregate based on hour (example of what this looks like, follows).  Note for multiple recordings within an hour, take the mean.  Then to test appropriate assumptions and forecast a week forward with confidence bands (80 and 95%). Add these to your existing files above - clearly labeled.  

## Pipes1 Forecast 

```{r}
#Insert Forecast Here 
```

## Pipes2 Forecast

```{r}
#Insert Forecast Here 
```

## R Script 

```{r, echo=T}
#Insert Script Here 
```

# Appendix 

## Part B 

### Model Summary 

**`ARIMA`:**
```{r}
summary(arima_auto)
```

**`STL - MNN`:**
```{r}
summary(stl_ndemp)
```

**`STL - MAdN`:**
```{r}
summary(stl_demp)
```

**`ets - MNM`:**
```{r}
summary(ets_model)
```

\newpage
### R Script 
```{r, echo=T, eval=F}
library(readxl)
library(tidyverse)
library(forecast)
library(imputeTS)
library(tsoutliers)

# load data
power_data <- read_excel("data/ResidentialCustomerForecastLoad-624.xlsx")

# Time Series
ts_data <- ts(power_data$KWH, frequency = 12, start = c(1998,1))

# Missing value imputation
ts_data <- na_interpolation(ts_data)

# STL decomposition
stl1 <- stl(ts_data, s.window = 'periodic')

# Handling outlier
outlier_func <- tsoutliers(ts_data, iterate = 2, lambda = "auto")

# Time Series - After outlier and imputation handeled
ts_data_o <- ts_data  # Let's treate outlier handled data seperatly for Modelling part.
ts_data_o[outlier_func$index] <- outlier_func$replacements

# Model#1: ARIMA
arima_auto <- auto.arima(ts_data_o)
arima_fc <- forecast(arima_auto, h=12)

# Model #2: STL (no-demped) - MNN
stl_ndemp <- stlf(ts_data_o, s.window = "periodic", robust=TRUE, h = 12)

# Model #2-2: STL (demped) - MAdN
stl_demp <- stlf(ts_data_o, damped=TRUE, s.window = "periodic", robust=TRUE, h = 12)

# Model #3: ets - MNM
ets_auto <- ets(ts_data_o)
ets_model <- forecast(ets_auto, h=12)

# tsCv - ARIMA -> it takes so much time. I got the results and saved them
##arima_cv <- function(x, h){forecast(Arima(x, order = c(3, 0, 2), seasonal = c(2, 1, 0), include.drift = TRUE), h=h)}
##e <- tsCV(ts_data_o, arima_cv, h=12)

# RMSEs -> tsCV takes lot of time to process so just saved the output
#rmse_train_arima <- arima_auto[2]
#rmse_test_arima <- sqrt(mean(e^2, na.rm=TRUE))

rmse_train_arima <- 589381.7
rmse_test_arima <- 725175

# Save output
write.csv(arima_fc, file="forecasts/POWER_ARIMA_FC.csv")
```
