---
title: 'DATA 624: Project 1 - Part B'
author: 'Sang Yoon (Andy) Hwang & Vinicio Haro'
date: 'October 22, 2019'
documentclass: book
subparagraph: yes
classoption: openany
output: 
  pdf_document:
    highlight: tango
    includes:
      in_header: preamble.tex
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: yes
    number_sections: true
    toc: yes
    toc_depth: 2
---

# Part B: Forecasting Power {-#part-b}

```{r library-B, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, comment=FALSE}
# processing
library(readxl)
#install.packages('tinytex')
library(tinytex)
#library(l3backend)
#install.packages('MiKTeX')
# graphs
library(ggplot2)

# formatting
library(default)
library(knitr)
library(kableExtra)
```

```{r settings-B, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, comment=FALSE}
# Set default augments for code chunks
knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, comment=F, fig.width=10, fig.height = 3)

# Set default augments for `kable_styling()` 
default(kable) <- list("latex")
default(kable_styling)  <- list(latex_options = c("hold_position", "striped"))

# Set default for ggplot theme
default(theme) <- list(axis.text.x = element_text(angle = 90, hjust = 1),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank(),
                       plot.title = element_text(color="#B85231", size=10, face="bold"),
                       plot.subtitle = (element_text(size=8, color="#000000")),
                       legend.title = (element_text(size=10, color="#000000", face="bold")),
                       strip.background = element_rect(color="#000000", 
                                                       fill="#F5E8E4", size=1, linetype="solid"),
                       strip.text.x = element_text(size = 8, color = "#000000", face="bold"))

# GGplot Palette
default(scale_color_brewer) <- list(palette = "OrRd")
```

> **Instructions:** Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.  Your assignment is to model these data and a monthly forecast for 2014.  The data is given in a single file.  The variable 'KWH' is power consumption in Kilowatt hours, the rest is straight forward.    Add these to your existing files above - clearly labeled.  

## Data Exploration and Processing {-#b-exploration}

Explore data. Process as needed. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
library(readxl)
library(forecast)
library(lubridate)
library(fpp2)
library(ggplot2)
library(forecast)
library(tseries)
library(imputeTS)
library(tsoutliers)
#install.packages('tsoutliers')
```

```{r warning=FALSE, message=FALSE}
#power_data <- read_excel("data/ResidentialCustomerForecastLoad-624.xlsx") 
library (readr)

power="https://raw.githubusercontent.com/vindication09/DATA-624/master/ResidentialCustomerForecastLoad-624.csv"

partb_data<-read_csv(url(power))

head(partb_data)
```

Transformed data into time-series with freq - 12. We are missing 2008 Sep data point.
```{r}
ts_data <- ts(partb_data$KWH, frequency = 12, start = c(1998,1))
ts_data
```

We impute missing data using TSImpute's interpolation method. 
```{r warning=FALSE, message=FALSE}
ts_data<-na.interpolation(ts_data)
```

Review the cycle of the time series to get an idea of the positions within the cycle. 
```{r warning=FALSE, message=FALSE}
cycle(ts_data)
```

Let's do quick EDA on ts_data. Outlier is detected - Min is significantly lower than median. We will handle this outlier. In general, data is fairly normally distributed as mean is not far from median.
```{r warning=FALSE, message=FALSE}
summary(ts_data)
hist(ts_data)
```

Let's do deeper EDA on ts_data and handle outlier using tsoutliers().
```{r warning=FALSE, message=FALSE}
#disable scientific notation (ONLY RUN ONCE)
options(scipen = 99999)

autoplot(ts_data) +
labs(title = "Monthly Residential Power Usage", subtitle = "01/98 - 12/13")+
theme_classic();

ggseasonplot(ts_data);

boxplot(ts_data~cycle(ts_data),xlab="Month", ylab = "Monthly Residential Power Usage");

ggsubseriesplot(ts_data);

stl(ts_data, s.window = 'periodic') %>% autoplot();

ggAcf(ts_data);
#Box.test(ts_data, type = c("Ljung-Box"))


# handling outlier
#fit <- nnetar(tsclean(ts_data))
outlier_func <- tsoutliers(ts_data, iterate = 2, lambda = "auto")
ts_data[outlier_func$index] <- outlier_func$replacements
```

Our initial plots reveal annual seasonality within this time series. The box plot/seasonality plot actually reveals where power consumption fluctuations occur within each of the cycke positions. We can speculate that this could be due to there being no major Holidays that require power draining decor plus we assume minimal AC usage during the cold months. 

We see power consumption increase between the months of June and August. This must be tied to AC usage during the warmer months of a year and finally power usage dips from September to Novemeber with a small spike in December. We speculate that thisis due to transitioning out of summer. The spike in December could be connected to the usage or Holiday lights being kept on. 

Within the overall TS plot, we see a dip in July 2010. This could be due to a power outtage during a hot summer month. This can certainly be considered to be an outlier within this TS. Using TSOutliers, we can actually identify the index where our outliers may be. TSoutliers also replaces the outlier using Box-Cox. If set lambda=auto, then TSoutliers will automatically perform Box-Cox transformation. 

The ACF plot shows that autocorrelations are well outside the significant space indicating the series is not white noise, non-stationary. 


## Data Model {-#b-model}

### Model #1: ARIMA
```{r warning=FALSE, message=FALSE}
arima_model <- auto.arima(ts_data)

arima_model <- forecast(arima_model, h=12)

autoplot(arima_model) + autolayer(fitted(arima_model))

checkresiduals(arima_model)
```


### Model #2: STL (no-demped) - MNN
```{r warning=FALSE, message=FALSE}
#stlf - etsmodel estimation --- M,N,N is chosen.
stl_ndemp <- stlf(ts_data, s.window = "periodic", robust=TRUE, h = 12)

# forecast plot
autoplot(stl_ndemp) + autolayer(fitted(stl_ndemp))

checkresiduals(stl_ndemp)
```

### Model #2-2: STL (demped) - MAdN
```{r warning=FALSE, message=FALSE}
#stlf - etsmodel estimation --- M, Ad, N is chosen.
stl_demp <- stlf(ts_data, damped=TRUE, s.window = "periodic", robust=TRUE, h = 12)

# forecast plot
autoplot(stl_demp) + autolayer(fitted(stl_demp))

checkresiduals(stl_demp)
```

### Model #3: ets - MNM
```{r warning=FALSE, message=FALSE}
# ETS models - MNM
ets_model <- ets(ts_data)

# forecast plot
autoplot(forecast(ets_model, h=12)) + autolayer(fitted(ets_model))

checkresiduals(ets_model)
```

### Model #4: Regression
```{r warning=FALSE, message=FALSE}
fit.reg <- tslm(ts_data ~ trend + season)
reg_model <- forecast(fit.reg, h = 12)

# forecast plot
autoplot(reg_model) + autolayer(fitted(reg_model))

checkresiduals(reg_model)
```

### Model #5: TBATS
```{r warning=FALSE, message=FALSE}
fit.tbats <- tbats(ts_data)
tbats_model <- forecast(fit.tbats, h = 12)

# forecast plot
autoplot(tbats_model) + autolayer(fitted(tbats_model))

checkresiduals(tbats_model)
```

### Accuracy of Models
```{r warning=FALSE, message=FALSE}
accuracy(arima_model);

accuracy(stl_ndemp);

accuracy(stl_demp);

accuracy(ets_model);

accuracy(reg_model);

accuracy(tbats_model)
```

 Out of the models we built, we can make some preliminary observations. The residuals for each of our models does not have a major deviance from normality, however Model #1: ARIMA residuals do not have an extended number of bins distorting the normality proximity but we can say it is still fairly normally distributed. 
 
 The residual ACF plots show residual autocorrelations for each of our models. Model #1: ARIMA has less autocorrelation than the other three models. Model 1 is well within the 95% limits indicated by the dotted blue lines. 
 
 If we examine the Ljung-Box test results for our models, the only model with a p-value > 0.05 is Model #1: ARIMA. This implies that the residuals from other models are not independent, hence not white noise. 
 
 In contrast, when we first attempted the analysis by building models without handling outlier (2010 Jul), the only model with a p-value < 0.05 was Model #3: ets - MNN and hence we accepted all the other models except for #3.
 
 Handling 1 outlier dramatically changed the outcome of Ljung-Box test.


## Forecast {-#b-forecast}

We will implement a cross validation method of testing for h=12. The process randomly chooses 12 points to measure and take the average of RMSEs. By definition, a lower RMSE on test set is attributed with a better forecast on unseen data. We only accepted Model #1 from Ljung-Box test and hence it is our final choice.

### Model #1: ARIMA
```{r warning=FALSE, message=FALSE}
arima_cv <- function(x, h){forecast(Arima(x, order = c(3, 0, 2), seasonal = c(2, 1, 0), include.drift = TRUE), h=h)}

e <- tsCV(ts_data, arima_cv, h=12)

sqrt(mean(e^2, na.rm=TRUE))
```

 Using Time series cross-validation, we compute RMSE on testset (h=12). We would have to pick the model with the lowest RMSE on test set as our final model if we had more than 1 model to compare. In our case, since we only have 1 model left after Ljung test, we  have no choice but to pick seasonal ARIMA model as our final choice. Cross-validation test shows that RMSE on test is around 720k when RMSE on training is around 589k. We can conclude the model is not necessarily overfitted. Given that MAPE on training is less than 7, it is not a suprising result.


## Discussion {-#b-discussion}

In our first phased implicit analysis, which is not recorded on this Rmd, where outlier (2010 - Jul) was not handled, STL - ANN was the best model in terms of RMSE on test set. As outlier was handled, on the other hand, STL models became invalid predictors as residuals were autocorrelated. We learned that handling outlier, even if it is just one data point, makes the result of modelling outcome completely.
