---
title: "DATA 624_Exercises_HA 6.2"
author: "Jeremy O'Brien"
date: "September 7, 2019"
output: html_document
---

```{r, message = FALSE, warning = FALSE}

if (!require('fpp2')) (install.packages('fpp2'))
if (!require('magrittr')) (install.packages('magrittr'))
if (!require('dplyr')) (install.packages('dplyr'))
if (!require('readxl')) (install.packages('readxl'))
if (!require('RCurl')) (install.packages('RCurl'))
if (!require('tsfeatures')) (install.packages('tsfeatures'))

```


###Exercise 6.9.2
####The _plastics_ dataset consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.

```{r, eval = FALSE}

help(plastics)
str(plastics)

```

<br>

####a. Plot the time series of sales of product A.  Can you identify seasonal fluctuations and/or a trend-cycle?

```{r, warning = FALSE}

# Set consistent formatting for plots
formatting <- theme(plot.title = element_text(hjust = .5)) + 
  theme(text = element_text(size = 10))
  
# Plot the monthly plastics time series together with moving averages of order 3, 5, 7, 9, 11, and 13
autoplot(plastics, main = 'Plastics Product A, Monthly Sales (Five Years)', series = 'Product A', size = .8) +
  autolayer(ma(plastics, order = 3), 
            series = '3-MA', 
            size = .7) +
  autolayer(ma(plastics, order = 5), 
            series = '5-MA', 
            size = .7) +
  autolayer(ma(plastics, order = 7), 
            series = '7-MA', 
            size = .7) +
  autolayer(ma(plastics, order = 9), 
            series = '9-MA', 
            size = .7) +
  autolayer(ma(plastics, order = 11), 
            series = '11-MA', 
            size = .7) +
  autolayer(ma(plastics, order = 13), 
            series = '13-MA', 
            size = .7) +
  formatting +
  ylab(label = '') +
  ylim(0, max(plastics)) +
  scale_color_manual(values = c('Product A' = 'black', 
                                '3-MA' = 'red', 
                                '5-MA' = 'rosybrown', 
                                '7-MA' = 'goldenrod3', 
                                '9-MA' = 'olivedrab', 
                                '11-MA' = 'royalblue4',
                                '13-MA' = 'midnightblue'),
                     breaks = c('Product A', 
                                '3-MA', 
                                '5-MA', 
                                '7-MA', 
                                '9-MA', 
                                '11-MA',
                                '13-MA')
                     )

```

####There's a clear seasonal fluctuation, exhibiting troughs at the beginning of each year and peaks over August through October.  Additionally, there's a consistent upward trend.

<br>

####b. Use a classical multiplicative decomposition to calculate the trend cycle and seasonal indices.

```{r}

# Produce multiplicative decomposition of the monthly plastics time series
plastics_decomp <- plastics %>% 
  decompose(type = 'multiplicative')

# Plot the multiplicative decomposition
plastics_decomp %>% 
  autoplot()

# Output the seasonal index
(function(x) print(paste(min(x), max(x))) ) (plastics_decomp$figure)

# The tsfeatures package provides statistics on strength of trend and seasonality, which doesn't seem to be a built-in feature of the forecast or other time series packages: https://rdrr.io/github/robjhyndman/tsfeatures/man/stl_features.html

# Output the trend strength and seasonal strength metrics 
plastics %>% 
  tsfeatures() %>%  
  select(trend, seasonal_strength) %>% 
  rename(trend_str = trend, season_str = seasonal_strength)

```

<br>

####c. Do the results support the graphical interpretations from part a?

####Yes, they do.  The multiplicative decomposition, with $F_S$ of .963 (close to the maximum of 1), substantiates the strong seasonal trend observed in the data.  The seaonsal index ranges between .71 and 1.23.  The trend is also strong, with $F_T$ of .92 (also close to the maximum of 1).

<br>

####d. Compute and plot the seasonally adjusted data.

```{r}

# Compute the seasonally adjusted monthly plastics time series
plastics_seasadj <- plastics / plastics_decomp$seasonal

# Plot the seasonally adjusted monthly plastics time series
autoplot(plastics_seasadj, 
         main = 'Plastics Product A, Seasonally Adjusted Monthly Sales (Five Years)', 
         series = 'Product A (Seasonally Adjusted)', 
         size = .7, 
         color = 'black') +
  formatting +
  ylab(label = '') +
  ylim(0, max(plastics_seasadj))

```

<br>

####e. Change one observation to be an outlier and recompute the seasonally adjusted data.  What is the effect of the outlier?

```{r}

# Artifically change single observation to create outlier between 1.5x IQR in middle of time series
plastics_2 <- plastics
plastics_2[25] <- quantile(plastics_2)[4] * 1.5
plastics_2_decomp <- decompose(plastics_2, type = 'multiplicative')
plastics_2_seasadj <- plastics_2 / plastics_2_decomp$seasonal

# Plot the seasonally adjusted monthly plastics time series
autoplot(plastics_2_seasadj, 
         main = 'Plastics Product A, Seasonally Adjusted Monthly Sales With Artificial Outlier (Five Years)', 
         series = 'Product A (Seasonally Adjusted)', 
         size = .7, 
         color = 'black') +
  formatting +
  ylab(label = '') +
  ylim(0, max(plastics_2_seasadj))

```

####The effect is to deflate the corresponding month in other years in the seasonally adjusted data.  In this case, the first month (January) of Year 3 was artifically inflated to approximately 2,111, which is 1.5 times larger than the upper bound of the inter-quartile range (the minimum value for an outlier).  The second months of years 1-2 and 4-5 have all fallen to compensate together for this outlier and now represent troughs.

<br>

####f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?

<br>

```{r}

# Create outliers at beginning and end of respectived time series
plastics_3 <- plastics
plastics_3[46] <- quantile(plastics_3)[4] * 1.5
plastics_3_decomp <- decompose(plastics_3, type = 'multiplicative')
plastics_3_seasadj <- plastics_3 / plastics_3_decomp$seasonal

plastics_4 <- plastics
plastics_4[3] <- quantile(plastics_4)[4] * 1.5
plastics_4_decomp <- decompose(plastics_4, type = 'multiplicative')
plastics_4_seasadj <- plastics_4 / plastics_4_decomp$seasonal

# autoplot(plastics_4_seasadj)

# Plot different outlier time series for comparison
autoplot(plastics_4_seasadj, 
         main = 'Plastics Product A, Seasonally Adjusted Monthly Sales (Five Years)', 
         series = 'SeasAdj', 
         size = .7) +
  autolayer(plastics_2_seasadj, 
            series = 'SeasAdj w/ Outlier: Yr3 Jan', 
            size = .7) +
  autolayer(plastics_3_seasadj, 
            series = 'SeasAdj w/ Outlier: Yr5 Aug', 
            size = .7) +
  autolayer(plastics_4_seasadj, 
            series = 'SeasAdj w/ Outlier: Yr1 Mar', 
            size = .7) +
  formatting +
  ylab(label = '') +
  ylim(0, max(c(
    max(plastics_seasadj),
    max(plastics_2_seasadj),
    max(plastics_3_seasadj), 
    max(plastics_4_seasadj)
    )
    )
    ) +
  scale_color_manual(values = c('SeasAdj' = 'black',
                                'SeasAdj w/ Outlier: Yr1 Mar' = 'red',
                                'SeasAdj w/ Outlier: Yr3 Jan' = 'goldenrod3',
                                'SeasAdj w/ Outlier: Yr5 Aug' = 'royalblue4'),
                     breaks = c('SeasAdj',
                                'SeasAdj w/ Outlier: Yr1 Mar',
                                'SeasAdj w/ Outlier: Yr3 Jan',
                                'SeasAdj w/ Outlier: Yr5 Aug'
                                )
  )

```

####We replaced values at the beginning (March of Year 1), middle (January of Year 3), and end (August of Year 5) of the time series with the same outlier value (1.5 times larger than the upper bound of the IQR) and then performing seasonal adjustment.  It appears that outliers at the start of the time series have the most impact on seasonally adjusted values with, larger spikes and compensatory troughs in other years.  These then decline as the time series continues, with smaller mdoerate impacts in the middle and smaller ones towards the end.  Our intuition was that outliers in the middle would exert the greatest impact, so this is a slightly puzzling finding.

***

<br>

```{r, echo = FALSE}

sessionInfo()

```